---
title: "Summer avgs"
author: "Abby Lewis"
date: "2022-11-22"
output: html_document
---

This file loads temp/DO data and calculates means during the late-summer period (July-Aug in the Northern Hemisphere). All code chunks must be run in order.

Table of contents:
Step 1: Load data and packages
Step 2: QAQC
Step 3: Filter to stratified period and calculate thermocline depths
Step 4: Calculate average values within each layer
Step 5: Add buoyancy frequency during the late-summer period
Step 6: Repeat steps 3, 4, and 5 for July 15-August 31
Step 7: Repeat steps 3, 4, and 5 for June/July/August

Step 1: Load data and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(rLakeAnalyzer)
source("thermo.depth.density.R")

## Load data and metadata from EDI
lat_long <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/1530/1/fadd3eaa25b5fdd1fc4efba70e660579")
#Load interpolated DO data file, created by "02 - Temp and DO interpolation.Rmd"
do <- read.csv("../Compiled data/temp_o2_interpolated.csv")
p <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/1530/1/0ece9d7b67cd49741ed7ee60192832e4")
```

Step 2: QAQC

```{r}
#Merge P, DO, and lake info (from lat long file)
full <- do %>%
  left_join(lat_long, by = c("LakeID")) %>%
  mutate(Date = as.Date(Date)) %>%
  filter(is.na(MaximumDepth_m) | !MaximumDepth_m < 6.4)

#Create a Date_22 column, so that I can filter to only the end of summer
full$Date_22 <- full$Date
year(full$Date_22) <- 2022

jpeg("../Figures/Data_hist.jpg", width = 3, height = 3, units = "in", res = 300)
full %>%
  group_by(Date, LakeID) %>%
  summarize(n = length(!is.na(Temp_C))) %>%
  filter(n > 0) %>%
  mutate(month = as.numeric(month(Date))) %>%
  ggplot(aes(x = month)) +
  geom_histogram(bins = 12) +
  scale_x_continuous(breaks = seq(1, 12, by = 2), labels = month.abb[seq(1, 12, by = 2)])+
  ylab("Total number of sampling events") +
  theme_bw() +
  theme(axis.title.x = element_blank())
dev.off()
```

Step 3: Filter to summer period and calculate thermocline depths

```{r}
full_trimmed <- full%>%
  filter(((Latitude_DD>0)&Date_22>=as.Date("2022-07-01")&Date_22<=as.Date("2022-08-31"))|
           ((Latitude_DD<0)&Date_22>=as.Date("2022-01-01")&Date_22<=as.Date("2022-02-28"))) %>%
  dplyr::select(-Date_22)

#Calculate thermocline depths
thermo_depths <- full_trimmed%>%
  group_by(Date, LakeID, Depth_m) %>%
  dplyr::summarize(Temp_C = mean(Temp_C, na.rm = T)) %>%
  filter(!is.na(Temp_C)) %>%
  ungroup() %>%
  group_by(Date, LakeID) %>%
  dplyr::summarize(epi_depth = meta.depths(Temp_C,Depth_m,mixed.cutoff = 0)[1],
            hypo_depth = meta.depths(Temp_C,Depth_m,mixed.cutoff = 0)[2],
            max_depth = max(Depth_m),
            thermo = thermo.depth.density(Temp_C,Depth_m, mixed.cutoff = 0.1, seasonal = F)) %>% #use custom density threshold function
  mutate(Year= year(Date),
         unstrat = as.numeric(is.na(thermo))) %>%
  group_by(Year,LakeID) %>%
  dplyr::summarize(epi_depth = mean(epi_depth, na.rm = T),
            hypo_depth = mean(hypo_depth, na.rm = T),
            count_unstrat = sum(unstrat),
            n = n())

#How many are removed by filtering out lakes with 10% of profiles being unstratified? 19
thermo_depths%>%
  group_by(LakeID) %>%
  dplyr::mutate(count_unstrat_tot = sum(count_unstrat),
                n = sum(n)) %>%
  filter((count_unstrat_tot/n) >=0.1) %>%
  ungroup() %>%
  summarize(lakes = length(unique(LakeID)))

#Remove years with unstratified profiles and lakes where 10% of years have unstratified profiles
thermo_depths_sum <- thermo_depths%>%
  group_by(LakeID) %>%
  dplyr::mutate(count_unstrat_tot = sum(count_unstrat),
                n = sum(n)) %>%
  filter((count_unstrat_tot/n) <0.1,
         count_unstrat == 0) %>%
  group_by(LakeID,Year) %>%
  dplyr::summarize(epi_sd = sd(epi_depth, na.rm = T),
            epi_depth = mean(epi_depth, na.rm = T),
            hypo_sd = sd(hypo_depth, na.rm = T),
            hypo_depth = mean(hypo_depth, na.rm = T))

#Export thermocline depths
write_csv(thermo_depths_sum, "../Compiled data/stratified_lakes.csv")
```

Step 4: Calculate average values within each layer

```{r}
#Add thermocline depths
full_with_thermo <- full_trimmed %>%
  mutate(Year = year(Date)) %>%
  full_join(thermo_depths_sum) %>%
  filter(!is.na(epi_depth),
         !is.na(hypo_depth),
         )
#Add discrete layer designations
summer_layers <- full_with_thermo %>%
  mutate(Layer = ifelse(!is.na(Depth_m) & Depth_m<epi_depth, 
                        "EPI", NA),
         Layer = ifelse(!is.na(Depth_m) & Depth_m>hypo_depth,
                        "HYPO", Layer),
         Layer = ifelse(!is.na(Depth_m) & Depth_m < hypo_depth & Depth_m > epi_depth,
                        "META",Layer)) %>%
  filter(!is.na(Layer))

#Calculate averages
summer_avgs <- summer_layers%>%
  group_by(LakeID,Year, Layer) %>% #not separating by measurement location. Is this a problem?
  dplyr::summarize(DO_mgL = mean(DO_mgL, na.rm = T),
                   DO_sat = mean(DO_sat, na.rm = T),
                   Temp_C = mean(Temp_C, na.rm = T))
```

Add surface and bottom that aren't defined by thermal layer

```{r}
#Include shallow lakes
full_shallow <- do %>%
  full_join(p) %>%
  left_join(lat_long, by = c("LakeID")) %>%
  mutate(Date = as.Date(Date)) %>%
  filter(is.na(MaximumDepth_m) | !MaximumDepth_m < 3)

full_shallow$Date_22 <- full_shallow$Date
year(full_shallow$Date_22) <- 2022

full_trimmed_shallow <- full_shallow %>%
  filter(((Latitude_DD>0) & Date_22 >= as.Date("2022-07-01") & Date_22 <= as.Date("2022-08-31"))|
           ((Latitude_DD<0) & Date_22 >= as.Date("2022-01-01") & Date_22 <= as.Date("2022-02-28"))) %>%
  dplyr::select(-Date_22)

surf_bot <- full_trimmed_shallow %>%
  group_by(Date, LakeID, Depth_m) %>%
  dplyr::summarize(Temp_C = mean(Temp_C, na.rm = T),
                   DO_mgL = mean(DO_mgL, na.rm = T),
                   TP_ugL = mean(TP_ugL, na.rm = T),
                   DOC_mgL = mean(DOC_mgL, na.rm = T),
                   Chla_ugL = mean(Chla_ugL, na.rm = T)) %>%
  filter(!is.na(Temp_C) | !is.na(DO_mgL)) %>%
  ungroup() %>%
  mutate(Year = year(Date)) %>%
  group_by(Year, LakeID) %>%
  dplyr::summarize(DO_mgL_SURF = mean(DO_mgL[Depth_m == 1], na.rm=T),
                   DO_mgL_BOT = mean(DO_mgL[Depth_m >= max(Depth_m)-1], na.rm=T),
                   Temp_C_SURF = mean(Temp_C[Depth_m == 1], na.rm=T),
                   Temp_C_BOT = mean(Temp_C[Depth_m >= max(Depth_m)-1], na.rm=T),
                   DOC_mgL_SURF = mean(DOC_mgL[Depth_m <= 1], na.rm=T),
                   Chla_ugL_SURF = mean(Chla_ugL[Depth_m <= 1], na.rm=T),
                   TP_ugL_SURF = mean(TP_ugL[Depth_m <= 1], na.rm=T))

buoyancy_shallow <- full_trimmed_shallow %>%
  group_by(Date, LakeID, Depth_m) %>%
  dplyr::summarize(Temp_C = mean(Temp_C, na.rm = T)) %>%
  filter(!is.na(Temp_C)) %>%
  ungroup() %>%
  group_by(Date, LakeID) %>%
  dplyr::summarize(buoyancy_freq = max(buoyancy.freq(Temp_C,Depth_m),na.rm=T)) %>%
  mutate(Year=year(Date)) %>%
  group_by(Year,LakeID) %>%
  dplyr::summarize(buoyancy_freq = mean(buoyancy_freq,na.rm=T))
```


Step 5: Add buoyancy frequency during the late-summer period

```{r}
#buoyancy <- full_trimmed%>%
#  group_by(Date, LakeID, Depth_m) %>%
#  dplyr::summarize(Temp_C = mean(Temp_C, na.rm = T)) %>%
#  filter(!is.na(Temp_C)) %>%
#  ungroup() %>%
#  group_by(Date, LakeID) %>%
#  dplyr::summarize(buoyancy_freq = max(buoyancy.freq(Temp_C,Depth_m),na.rm=T)) %>%
#  mutate(Year=year(Date)) %>%
#  group_by(Year,LakeID) %>%
#  dplyr::summarize(buoyancy_freq = mean(buoyancy_freq,na.rm=T))

summer_avgs <- surf_bot %>%
  full_join(summer_avgs) %>%
  left_join(buoyancy_shallow) 

#All done!
write.csv(summer_avgs, "../Compiled data/summer_averages_wi.csv", row.names = F)
```

Step 6: Repeat steps 3, 4, and 5 for July 15-August 31

```{r}
full_trimmed <- full%>%
  filter(((Latitude_DD>0)&Date_22>=as.Date("2022-07-15")&Date_22<=as.Date("2022-08-31"))|
           ((Latitude_DD<0)&Date_22>=as.Date("2022-01-15")&Date_22<=as.Date("2022-02-28"))) %>%
  dplyr::select(-Date_22) 

#Calculate thermocline depths
thermo_depths <- full_trimmed%>%
  group_by(Date, LakeID, Depth_m) %>%
  dplyr::summarize(Temp_C = mean(Temp_C, na.rm = T)) %>%
  filter(!is.na(Temp_C)) %>%
  ungroup() %>%
  group_by(Date, LakeID) %>%
  dplyr::summarize(epi_depth = meta.depths(Temp_C,Depth_m,mixed.cutoff = 0)[1],
            hypo_depth = meta.depths(Temp_C,Depth_m,mixed.cutoff = 0)[2],
            max_depth = max(Depth_m),
            thermo = thermo.depth.density(Temp_C,Depth_m, mixed.cutoff = 0.1, seasonal = F)) %>% #use custom density threshold function
  mutate(Year= year(Date),
         unstrat = as.numeric(is.na(thermo))) %>%
  group_by(Year,LakeID) %>%
  dplyr::summarize(epi_depth = mean(epi_depth, na.rm = T),
            hypo_depth = mean(hypo_depth, na.rm = T),
            count_unstrat = sum(unstrat),
            n = n())

#Remove years with unstratified profiles and lakes where 10% of years have unstratified profiles
thermo_depths_sum <- thermo_depths%>%
  group_by(LakeID) %>%
  dplyr::mutate(count_unstrat_tot = sum(count_unstrat),
                n = sum(n)) %>%
  filter((count_unstrat_tot/n) <0.1,
         count_unstrat == 0) %>%
  group_by(LakeID,Year) %>%
  dplyr::summarize(epi_sd = sd(epi_depth, na.rm = T),
            epi_depth = mean(epi_depth, na.rm = T),
            hypo_sd = sd(hypo_depth, na.rm = T),
            hypo_depth = mean(hypo_depth, na.rm = T))

#Add thermocline depths
full_with_thermo <- full_trimmed %>%
  mutate(Year = year(Date)) %>%
  full_join(thermo_depths_sum) %>%
  filter(!is.na(epi_depth),
         !is.na(hypo_depth),
         )
#Add discrete layer designations
summer_layers <- full_with_thermo %>%
  mutate(Layer = ifelse(!is.na(Depth_m) & Depth_m<epi_depth, 
                        "EPI", NA),
         Layer = ifelse(!is.na(Depth_m) & Depth_m>hypo_depth,
                        "HYPO", Layer),
         Layer = ifelse(!is.na(Depth_m) & Depth_m < hypo_depth & Depth_m > epi_depth,
                        "META",Layer)) %>%
  filter(!is.na(Layer))

#Calculate averages
summer_avgs <- summer_layers%>%
  group_by(LakeID,Year, Layer) %>% #not separating by measurement location. Is this a problem?
  dplyr::summarize(DO_mgL = mean(DO_mgL, na.rm = T),
                   DO_sat = mean(DO_sat, na.rm = T),
                   Temp_C = mean(Temp_C, na.rm = T))

buoyancy <- full_trimmed%>%
  group_by(Date, LakeID, Depth_m) %>%
  dplyr::summarize(Temp_C = mean(Temp_C, na.rm = T)) %>%
  filter(!is.na(Temp_C)) %>%
  ungroup() %>%
  group_by(Date, LakeID) %>%
  dplyr::summarize(buoyancy_freq = max(buoyancy.freq(Temp_C,Depth_m),na.rm=T)) %>%
  mutate(Year=year(Date)) %>%
  group_by(Year,LakeID) %>%
  dplyr::summarize(buoyancy_freq = mean(buoyancy_freq,na.rm=T))

summer_avgs <- summer_avgs%>%
  left_join(buoyancy)

#All done!
write.csv(summer_avgs, "../Compiled data/summer_averages_wi_15th_to_31st.csv", row.names = F)
```

Step 7: Repeat steps 3, 4, and 5 for June/July/August

```{r}
full_date <- full %>%
  mutate(Date = ifelse(Latitude_DD < 0, Date + months(6), Date)) %>%
  mutate(Date = as.Date(Date, origin = "1970-01-01"),
         Date_22 = Date)

year(full_date$Date_22) <- 2022

full_trimmed <- full_date %>%
  filter(Date_22 >= "2022-06-01", 
         Date_22 <= "2022-08-31") %>%
  dplyr::select(-Date_22)

#Calculate thermocline depths
thermo_depths <- full_trimmed%>%
  group_by(Date, LakeID, Depth_m) %>%
  dplyr::summarize(Temp_C = mean(Temp_C, na.rm = T)) %>%
  filter(!is.na(Temp_C)) %>%
  ungroup() %>%
  group_by(Date, LakeID) %>%
  dplyr::summarize(epi_depth = meta.depths(Temp_C,Depth_m,mixed.cutoff = 0)[1],
            hypo_depth = meta.depths(Temp_C,Depth_m,mixed.cutoff = 0)[2],
            max_depth = max(Depth_m),
            thermo = thermo.depth.density(Temp_C,Depth_m, mixed.cutoff = 0.1, seasonal = F)) %>% #use custom density threshold function
  mutate(Year = year(Date),
         unstrat = as.numeric(is.na(thermo))) %>%
  group_by(Year,LakeID) %>%
  dplyr::summarize(epi_depth = mean(epi_depth, na.rm = T),
            hypo_depth = mean(hypo_depth, na.rm = T),
            count_unstrat = sum(unstrat),
            n = n())

#Remove years with unstratified profiles and lakes where 10% of years have unstratified profiles
thermo_depths_sum <- thermo_depths%>%
  group_by(LakeID) %>%
  dplyr::mutate(count_unstrat_tot = sum(count_unstrat),
                n = sum(n)) %>%
  filter((count_unstrat_tot/n) < 0.1,
         count_unstrat == 0) %>%
  group_by(LakeID,Year) %>%
  dplyr::summarize(epi_sd = sd(epi_depth, na.rm = T),
            epi_depth = mean(epi_depth, na.rm = T),
            hypo_sd = sd(hypo_depth, na.rm = T),
            hypo_depth = mean(hypo_depth, na.rm = T))

#Add thermocline depths
full_with_thermo <- full_trimmed %>%
  mutate(Year = year(Date)) %>%
  full_join(thermo_depths_sum) %>%
  filter(!is.na(epi_depth),
         !is.na(hypo_depth),
         )
#Add discrete layer designations
summer_layers <- full_with_thermo %>%
  mutate(Layer = ifelse(!is.na(Depth_m) & Depth_m<epi_depth, 
                        "EPI", NA),
         Layer = ifelse(!is.na(Depth_m) & Depth_m>hypo_depth,
                        "HYPO", Layer),
         Layer = ifelse(!is.na(Depth_m) & Depth_m < hypo_depth & Depth_m > epi_depth,
                        "META",Layer)) %>%
  filter(!is.na(Layer))

#Calculate averages
summer_avgs <- summer_layers%>%
  group_by(LakeID,Year, Layer) %>% #not separating by measurement location. Is this a problem?
  dplyr::summarize(DO_mgL = mean(DO_mgL, na.rm = T),
                   DO_sat = mean(DO_sat, na.rm = T),
                   Temp_C = mean(Temp_C, na.rm = T))

buoyancy <- full_trimmed%>%
  group_by(Date, LakeID, Depth_m) %>%
  dplyr::summarize(Temp_C = mean(Temp_C, na.rm = T)) %>%
  filter(!is.na(Temp_C)) %>%
  ungroup() %>%
  group_by(Date, LakeID) %>%
  dplyr::summarize(buoyancy_freq = max(buoyancy.freq(Temp_C,Depth_m),na.rm=T)) %>%
  mutate(Year=year(Date)) %>%
  group_by(Year,LakeID) %>%
  dplyr::summarize(buoyancy_freq = mean(buoyancy_freq,na.rm=T))

summer_avgs <- summer_avgs%>%
  left_join(buoyancy)

#All done!
write.csv(summer_avgs, "../Compiled data/summer_averages_wi_JJA.csv", row.names = F)
```