---
title: "Summer avgs"
author: "Abby Lewis"
date: "2022-11-22"
output: html_document
---

This file loads temp/DO/productivity data and calculates mean concentrations of each focal variable during the late-summer period (July 15-Aug 31 in the Northern Hemisphere). Functionally, all of the code here will be very similar to "03 - Stratified avgs.Rmd", except that this file is ONLY for the late-summer period, whereas that file is for the whole stratified period. All code chunks must be run in order.

Table of contents:
Step 1: Load data and packages
Step 2: QAQC
Step 3: Filter to stratified period and calculate thermocline depths
Step 4: Calculate average values within each layer
Step 5: Add buoyancy frequency during the late-summer period


Step 1: Load data and packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(rLakeAnalyzer)
library(openair)
source("thermo.depth.density.R")

## Load productivity data and metadata from EDI
p <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/1530/1/0ece9d7b67cd49741ed7ee60192832e4")
lat_long <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/1530/1/fadd3eaa25b5fdd1fc4efba70e660579")
#Load interpolated DO data file, created by "02 - Temp and DO interpolation.Rmd"
do <- read.csv("../Compiled data/temp_o2_interpolated.csv")
```


Step 2: QAQC
```{r}
#Merge P, DO, and lake info (from lat long file)
full <- p%>%
  full_join(do, by = c("Date","LakeID","Depth_m"))%>%
  full_join(lat_long, by = c("LakeID"))%>%
  mutate(Date = as.Date(Date))

#Create a Date_22 column, so that I can filter to only the end of summer
full$Date_22 <- full$Date
year(full$Date_22) <-2022

#There are a few rows with missing dates because these lakes are in the lat long database and not do/p
full%>%
  filter(is.na(as.Date(Date)))

#Remove these NAs for now
full <- full%>%
  filter(!is.na(Date))

#Missing depths
missing_depths <- full%>%
  filter(is.na(Depth_m)) #Fair # of missing depths, mostly from the Jane et al. data publication
unique(missing_depths$LakeID)

missing_depth_and_interval <- full%>%
  filter(is.na(Depth_m)&is.na(Interval)) #Most of the missing depths that are NOT from the Jane et al. data publication have the interval filled in
unique(missing_depth_and_interval$LakeID)
```


Step 3: Filter to summer period
```{r}
full_trimmed <- full%>%
  filter(((Latitude_DD>0)&Date_22>=as.Date("2022-07-01")&Date_22<=as.Date("2022-08-31"))|
           ((Latitude_DD<0)&Date_22>=as.Date("2022-01-01")&Date_22<=as.Date("2022-02-28")))%>%
  dplyr::select(-Date_22)%>%
  mutate(start_date = ifelse(Latitude_DD>0,"2022-07-01","2022-01-01"))
```


Step 4: Calculate average values within each layer
```{r}
#Add thermocline depths
full_surface <- full_trimmed%>%
  mutate(Year = year(Date))%>%
  filter(!is.na(Depth_m)&Depth_m<2 | !is.na(Interval)&Interval=="EPILIMNION") %>%
  mutate(Layer = "SURF") %>%
  dplyr::select(-X)

full_bottom <- full_trimmed%>%
  mutate(Year = year(Date))%>%
  group_by(LakeID)%>%
  mutate(max_depth = max(Depth_m, na.rm = T),
         Layer = ifelse(Depth_m > (max_depth - 0.2*max_depth), "BOT", NA)) %>%
  filter(is.na(max_depth) | max_depth >= 2,
         abs(max_depth-MaximumDepth_m) < 0.3 * MaximumDepth_m | 
           abs(max_depth-MaximumDepth_m) <= 2
         )%>%
  filter(!is.na(Layer)) %>%
  dplyr::select(-max_depth, -X)

#Calculate averages
summer_avgs <- full_surface %>%
  full_join(full_bottom) %>%
  mutate(start_date = as.Date(paste0(year(Date),"-",month(start_date),"-",day(start_date))))%>%
  group_by(LakeID, Year, Layer) %>%
  dplyr::summarize(TP_ugL = mean(TP_ugL, na.rm = T),
                   TP_date = as.numeric(mean(Date[!is.na(TP_ugL)],na.rm=T)-unique(start_date)),
                   DOC_mgL = mean(DOC_mgL, na.rm = T),
                   DOC_date = as.numeric(mean(Date[!is.na(DOC_mgL)],na.rm=T)-unique(start_date)),
                   DO_mgL = mean(DO_mgL, na.rm = T),
                   DO_date = as.numeric(mean(Date[!is.na(DO_mgL)],na.rm=T)-unique(start_date)),
                   Chla_ugL = mean(Chla_ugL, na.rm = T),
                   Chla_date = as.numeric(mean(Date[!is.na(DO_mgL)],na.rm=T)-unique(start_date)),
                   Temp_C = mean(Temp_C, na.rm = T),
                   Temp_date = as.numeric(mean(Date[!is.na(Temp_C)],na.rm=T)-unique(start_date)),
                   TN_ugL = mean(TN_ugL, na.rm = T),
                   TN_date = as.numeric(mean(Date[!is.na(TN_ugL)],na.rm=T)-unique(start_date)))
```

Step 5: Add buoyancy frequency during the late-summer period
```{r}
buoyancy <- full_trimmed%>%
  group_by(Date, LakeID, Depth_m)%>%
  dplyr::summarize(Temp_C = mean(Temp_C, na.rm = T))%>%
  filter(!is.na(Temp_C))%>%
  ungroup()%>%
  group_by(Date, LakeID)%>%
  dplyr::summarize(buoyancy_freq = max(buoyancy.freq(Temp_C,Depth_m),na.rm=T))%>%
  mutate(Year=year(Date))%>%
  group_by(Year,LakeID)%>%
  dplyr::summarize(buoyancy_freq = mean(buoyancy_freq,na.rm=T))

summer_avgs <- summer_avgs%>%
  left_join(buoyancy)

#All done!
write.csv(summer_avgs, "../Compiled data/summer_averages_wi_surface.csv", row.names = F)
```
