---
title: "Calculate correlations"
author: "Abby Lewis"
date: "2023-05-22"
output: html_document
---

This file calculates the correlation between mean air temperature and surface- and bottom-water temperature and dissolved oxygen.

Table of contents:

- Step 1: Load packages and data
- Step 2: Load and format air temperature data
- Step 3: Calculate and export 30-day rolling mean correlations
- Step 4: Try using Spearman's correlation, rather than partial Spearman's correlation
- Step 5: Re-run with 10-day window
- Step 6: Re-run with 50-day window
- Step 7: Re-run for July 15th to August 31st
- Step 8: Re-run for June, July and August

Step 1: Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(rMR)
source("correlations_doy.R")

# Load saved data
with_temp = read.csv("../Compiled data/All_data_annual.csv")
lat_long <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/1530/1/fadd3eaa25b5fdd1fc4efba70e660579")

# Make a directory for Figures if there isn't one
if (!dir.exists("../Figures")){dir.create("../Figures")}
```

Step 2: Load and format air temperature data

```{r}
daily_met_raw <- read.csv("../Compiled data/historical_met_output_era5_daily.csv")
daily_met <- daily_met_raw %>%
  mutate(Date = as.Date(Date),
         Date = ifelse(Lat < 0, Date - months(6), Date),
         Date = as.Date(Date, origin = "1970-01-01")) %>%
  dplyr::select(-Lat, -Lon) %>%
  mutate(Year = year(Date)) %>%
  filter(Year <= 2022)
rm(daily_met_raw)
```

Step 3: Calculate and export rolling means (30-day)

```{r}
many_lake_temp_sum <- with_temp %>%
  left_join(daily_met) %>%
  mutate(doy = yday(Date),
         Year = ifelse(month(Date) >= 9, Year + 1, Year)) %>%
  group_by(LakeID) %>%
  arrange(Date) %>%
  mutate(Temp_C = zoo::rollmean(Temp_C, 30, align = "right", fill = NA),
         Wind_ms = zoo::rollmean(Wind_ms, 30, align = "right", fill = NA),
         Shortwave_mJm2 = zoo::rollmean(Shortwave_mJm2, 30, align = "right", fill = NA))

# ---
calc_corrs <- function(driver){
  hypo_do_roll <- many_lake_temp_sum_small %>%
    mutate(hypoxic = ifelse(max(DO_mgL_HYPO, na.rm = T) < 1,"yes","no")) %>%
    filter(hypoxic=="no") %>%
    correlations_doy("DO_mgL_HYPO", driver, "hypolimnetic DO")
  
  hypo_demand_roll <- correlations_doy(many_lake_temp_sum_small, 
                                       "DO_demand_mgLd_HYPO", driver, 
                                       "Summer VHOD ")
  
  epi_temp_roll <- correlations_doy(many_lake_temp_sum_small, 
                                    "Temp_C_EPI", driver, 
                                    "Summer epi. temperature ")
  
  hypo_temp_roll <- correlations_doy(many_lake_temp_sum_small, 
                                     "Temp_C_HYPO", driver, 
                                     "Summer hypo. temperature ")
  
  all <- list(
    hypo_do_roll %>%
      mutate(var = "Summer hypo. DO"),
    epi_temp_roll %>%
      mutate(var = "Summer epi. temperature"),
    hypo_temp_roll %>%
      mutate(var = "Summer hypo. temperature"),
    hypo_demand_roll %>% 
      mutate(var = "Summer VHOD"))
  
  all <- data.table::rbindlist(all) %>% 
    mutate(driver = driver)
  
  return(all)
}

drivers <- c("Temp_C", "Wind_ms", "Shortwave_mJm2")

many_lake_temp_sum_small <- many_lake_temp_sum %>%
  dplyr::select(LakeID, Date, doy, Year,
                Temp_C, Wind_ms, Shortwave_mJm2, 
                DO_mgL_HYPO, DO_demand_mgLd_HYPO, Temp_C_EPI, Temp_C_HYPO)

plan("multisession")
all_drivers <- furrr::future_map_dfr(drivers, 
                                     calc_corrs, 
                                     .progress = T)

write.csv(all_drivers, "../Compiled data/Correlations - all met.csv", row.names = F)
```