---
title: "Air temp analysis"
author: "Abby Lewis"
date: "2023-05-22"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

This file calculates "seasonal ecological memory" as the maximum absolute correlation between our summer focal variables and 30-day rolling mean air temperature during July--August (summer) minus the maximum correlation during spring (March-May).

Table of contents:

-   Step 1: Load packages and data
-   Step 2: Calculate seasonal ecological memory
-   Step 3: Generate Figure 1 (summary boxplot)
-   Step 4: Plot rolling means (hypolimnetic variables)
-   Step 5: Identify most influential dates
-   Step 6: Plot multiple rolling-mean windows (SI)

Step 1: Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)

#Set global options
sp <- yday(c("2022-03-01", "2022-05-31"))
su <- yday(c("2022-07-01", "2022-08-31"))
alpha <- 0.00001

#Load saved data
lat_long <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/1530/1/fadd3eaa25b5fdd1fc4efba70e660579")
all <- read.csv("../Compiled data/Correlations - 30 day rolling mean.csv") %>%
  mutate(sig = ifelse(p < alpha, T, F))
onset <- read.csv("../Compiled data/stratification_onset.csv")
hypo_do_sat <- read.csv("../Compiled data/Correlations - 30 day rolling mean - DO sat.csv")

# Make a directory for Figures if there isn't one
if (!dir.exists("../Figures")){dir.create("../Figures")}
```

Step 2: Calculate seasonal ecological memory

```{r}
memory <- all %>% 
  filter(var %in% c("Summer epi. temperature",
                    "Summer hypo. temperature",
                    "Summer VHOD",
                    "Summer hypo. DO")) %>%
  mutate(var = factor(var, levels = c("Summer epi. temperature",
                                      "Summer hypo. temperature",
                                      "Summer hypo. DO",
                                      "Summer VHOD"),
                      labels = c("Summer\nepi. temperature",
                                 "Summer\nhypo. temperature",
                                 "Summer\nhypo. DO",
                                 "Summer\nVHOD")),
         season = ifelse(doy >= sp[1] & doy <= sp[2], "spring", 
                         ifelse(doy >= su[1] & doy <= su[2], "summer", NA))) %>%
  group_by(var) %>%
  mutate(p = wilcox.test(monthly_correlation)$p.value) %>%
  filter(!is.na(season)) %>%
  group_by(LakeID, var, season, p) %>%
  summarize(max = max(abs(monthly_correlation))) %>%
  pivot_wider(names_from = season, values_from = max) %>%
  mutate(M = (summer - spring)) 

memory_sat <- hypo_do_sat %>%
  mutate(season = ifelse(doy >= sp[1] & doy <= sp[2], "spring", 
                         ifelse(doy >= su[1] & doy <= su[2], "summer", NA))) %>%
  group_by(var) %>%
  mutate(p = wilcox.test(monthly_correlation)$p.value) %>%
  filter(!is.na(season)) %>%
  group_by(LakeID, var, season, p) %>%
  summarize(max = max(abs(monthly_correlation))) %>%
  pivot_wider(names_from = season, values_from = max) %>%
  mutate(M = (summer - spring)) 
```

Step 3: Generate Figure 1 (summary boxplot)

```{r}
jpeg("../Figures/Monthly correlations - sum long concept v2.jpg", width = 3.5, height = 3, res = 600, units = "in")
memory %>%
  filter(!var == "Summer\nepi. DO sat") %>%
  group_by(var) %>%
  mutate(n = length(unique(LakeID)),
         var = paste0(var, " (n = ",n,")")) %>%
  mutate(var = factor(var, 
                      levels = c("Summer\nepi. temperature (n = 438)",
                                 "Summer\nhypo. temperature (n = 399)",
                                 "Summer\nVHOD (n = 194)",
                                 "Summer\nhypo. DO (n = 322)"),
                      labels = c("Surface-water\ntemperature (n = 438)",
                                 "Bottom-water\ntemperature (n = 399)",
                                 "Bottom-water oxygen\ndemand  (n = 194)",
                                 "Bottom-water dissolved\noxygen (n = 322)")))%>%
  group_by(var) %>%
  mutate(color = ifelse(p > alpha, "Insignificant", 
                        ifelse(median(M) > 0, "median > 0", "median < 0")),
         color = factor(color, levels = c("median < 0", 
                                          "Insignificant", 
                                          "median > 0"))) %>%
  ggplot(aes(y = var, x = M, color = color))+
  geom_vline(xintercept = 0, color = "grey70")+
  geom_boxplot()+
  annotate("text", x = -0.7, y = 5, label = "More correlated\nwith spring", 
           lineheight = .9, size = 3, hjust = 0, color = "#25469a")+
  annotate("text", x = 0.7, y = 5, label = "More correlated\nwith summer", 
           lineheight = .9, size = 3, hjust = 1, color = "#c16586")+
  theme_bw() +
  ggtitle("\n")+
  scale_y_discrete(limits=rev, position = "right") +
  xlab("Relative correlation with summer\nvs. spring air temperature")+
  guides(color = guide_legend(nrow = 2))+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(size= 8),
        legend.position = "none",
        legend.text = element_text(size = 8),
        legend.key.size = unit(1, "line"),
        legend.title = element_text(size = 8),
        legend.margin=margin(c(0,0,0,0)),
        legend.box.margin=margin(t = -2,l = 0, r = 0, b = -5),
        #axis.text.y=element_text(colour=c("#25469a","#25469a","#25469a","#c16586"))
        )+
  scale_color_manual(values = c("#25469a","#c16586"), 
                     name = paste0("Boxplot color\n(p < ", alpha, ")"))+
  coord_cartesian(xlim = c(-0.65,0.65), ylim = c(1,4), clip = "off")
dev.off()
```

Step 3b: Compare hypo DO concentration (mg/L) to hypo DO saturation (%)

```{r}
jpeg("../Figures/Monthly correlations - DO sat comp.jpg", width = 3.5, height = 2, res = 600, units = "in")
memory_sat %>%
  mutate(var = "Summer hypo. DO\nsaturation") %>%
  full_join(memory %>% 
              filter(var == "Summer\nhypo. DO") %>%
              mutate(var = "Summer hypo. DO\nconcentration")) %>%
  group_by(var) %>%
  mutate(n = length(unique(LakeID)),
         var = paste0(var, " (n = ",n,")")) %>%
  group_by(var) %>%
  mutate(color = ifelse(p > alpha, "Insignificant", 
                        ifelse(median(M) > 0, "median > 0", "median < 0")),
         color = factor(color, levels = c("median < 0", 
                                          "Insignificant", 
                                          "median > 0"))) %>%
  ggplot(aes(y = var, x = M, color = color))+
  geom_vline(xintercept = 0, color = "grey70")+
  geom_boxplot()+
  annotate("text", x = -0.7, y = 3, label = "More correlated\nwith spring", 
           lineheight = .9, size = 3, hjust = 0, color = "#25469a")+
  annotate("text", x = 0.7, y = 3, label = "More correlated\nwith summer", 
           lineheight = .9, size = 3, hjust = 1, color = "#c16586")+
  theme_bw() +
  ggtitle("\n")+
  scale_y_discrete(limits=rev, position = "right") +
  xlab("Relative correlation with summer\nvs. spring air temperature")+
  guides(color = guide_legend(nrow = 2))+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(size= 8),
        legend.position = "none",
        legend.text = element_text(size = 8),
        legend.key.size = unit(1, "line"),
        legend.title = element_text(size = 8),
        legend.margin=margin(c(0,0,0,0)),
        legend.box.margin=margin(t = -2,l = 0, r = 0, b = -5),
        #axis.text.y=element_text(colour=c("#25469a","#25469a","#25469a","#c16586"))
        )+
  scale_color_manual(values = c("#25469a","#c16586"), 
                     name = paste0("Boxplot color\n(p < ", alpha, ")"))+
  coord_cartesian(xlim = c(-0.65,0.65), ylim = c(1,2), clip = "off")
dev.off()
```

Step 4: Plot rolling means (hypolimnetic variables)

```{r}
#Function to deliniate periods of significant vs insignificant correlations
group_fun <- function(sig) {
  group <- rep(1, length(sig))
  for (i in 2:length(sig)) {
    if(sig[i] == sig[i - 1]) {
      group[i] <- group[i - 1]
    } else {
      group[i] <- group[i - 1] + 1
    }
  }
  return(group)
}

#Calculate mean and SD
all_plot <- all %>%
  group_by(var) %>%
  mutate(n = max(n), #Some lakes are missing from January because our air temp only goes back to 1980
         var = paste0(var, " (n = ",n,")")) %>%
  group_by(var, doy, sig) %>%
  summarise(mean = mean(monthly_correlation, na.rm = T),
            sd = sd(monthly_correlation, na.rm = T)) 

#Use group function
all_plot$group <- 1
for(i in 1:length(unique(all_plot$var))){
  all_plot$group[all_plot$var == unique(all_plot$var)[i]] <- 
    group_fun(all_plot$sig[all_plot$var == unique(all_plot$var)[i]])
}

#To get smooth ribbons, we need to manually duplicate datapoints
all_plot1 <- all_plot %>%
  mutate(DateTime = as.Date("2022-01-01") + days(doy))
all_plot2 <- all_plot %>%
  ungroup() %>%
  mutate(DateTime = as.Date("2022-01-01") + days(doy) + hours(23) + minutes(59),
         mean = ifelse(!is.na(lead(var)) & lead(var) == var, lead(mean), mean),
         sd = ifelse(lead(var) == var, lead(sd), sd))

#Plot hypolimnetic variables
hypo_vars <- all_plot1 %>%
  full_join(all_plot2) %>%
  filter(doy>=30) %>% #remove first 29 days (these include previous year)
  filter(var %in% c("Summer hypo. temperature (n = 399)",
                    "Summer VHOD (n = 194)",
                    "Summer hypo. DO (n = 322)")) %>%
  mutate(var = factor(var, 
                      levels = c("Summer hypo. temperature (n = 399)",
                                 "Summer VHOD (n = 194)",
                                 "Summer hypo. DO (n = 322)"),
                      labels = c("Summer bottom-water\ntemperature (n = 399)",
                                 "Summer bottom-water\noxygen demand (n = 194)",
                                 "Summer bottom-water\ndissolved oxygen (n = 322)")),
         layer = ifelse(grepl("epi", var), "epi", "hypo")) %>%
  ggplot() + 
  geom_hline(yintercept=0)+
  geom_vline(xintercept = as_datetime("2022-01-01")+
               days(median(onset$Onset, na.rm = T)),
             color = "grey50", lty = "dashed")+
  geom_ribbon(aes(x = DateTime, 
                  ymin = mean - sd, ymax = mean + sd,
                  fill = sig, group = group),
              alpha = 0.5) +
  geom_line(aes(x = DateTime, y = mean)) +
  xlab("Air temperature date (right-aligned 30-day rolling mean)")+
  ylab("Correlation")+
  theme_bw()+
  scale_fill_manual(values = c("grey40","#0FA9E6"))+
  labs(fill = paste0("p < ",alpha)) +
  facet_wrap(~var, nrow = 1) +
  theme(legend.position = "bottom",
        legend.margin=margin(c(0,0,0,0)),
        strip.background = element_blank())

#Save plot
jpeg("../Figures/Monthly correlations - hypo daily roll mean sd.jpg", 
     width = 6, height = 2.5, res = 300, units = "in")
hypo_vars
dev.off()
```

Step 5: Identify most influential dates

What dates are most correlated within the windows we assigned? How do these relate to stratification onset?

```{r}
dates_to_plot <- all %>%
  mutate(season = ifelse(doy >= sp[1] & doy <= sp[2], "spring", 
                         ifelse(doy >= su[1] & doy <= su[2], "summer", NA))) %>%
  filter(!is.na(season)) %>%
  group_by(var, LakeID, season) %>%
  summarize(max_doy = doy[which.max(abs(monthly_correlation))]) 

dates_to_plot %>%
  left_join(lat_long) %>%
  #filter(abs(Latitude_DD) > 30) %>%
  ggplot(aes(x = as.Date("2022-01-01") + max_doy))+
  geom_histogram(binwidth = 2)+
  xlab("Date of strongest correlation")+
  theme_bw() +
  facet_grid(cols = vars(season), rows = vars(var), scales = "free_x")

date_start <- read_csv("../Compiled data/stratification_onset.csv")

dates <- date_start %>%
  group_by(LakeID) %>%
  summarize(mean_start = mean(Onset, na.rm = T)) %>%
  filter(!is.na(mean_start))

dates_to_plot %>%
  filter(season == "spring",
         var == "Summer hypo. temperature") %>%
  inner_join(dates) %>%
  ggplot(aes(x = mean_start, y = max_doy, group = LakeID)) +
  geom_abline(slope = 1) +
  geom_smooth(method = "lm")+
  geom_point()

dates_export <- dates_to_plot %>%
  ungroup() %>%
  filter(grepl("temperature", var)) %>%
  filter((grepl("hypo", var) & season=="summer") | #epi
           (grepl("hypo", var) & season=="spring")) %>%
  dplyr::select(-var) 

write.csv(dates_export, 
          "../Compiled data/strongest_correlation_dates.csv", 
          row.names = F)

dates_export_both <- dates_to_plot %>%
  ungroup() %>%
  filter(grepl("temperature", var))

write.csv(dates_export_both, 
          "../Compiled data/strongest_correlation_dates_both.csv", 
          row.names = F)

dates_export_hypo <- dates_to_plot %>%
  ungroup() %>%
  filter(grepl("hypo", var)) %>%
  filter((grepl("temperature", var)) |
           (grepl("DO", var))) 

write.csv(dates_export_hypo, 
          "../Compiled data/strongest_correlation_dates_hypo.csv", 
          row.names = F)
```

Step 6: Plot multiple rolling-mean windows (SI)

```{r}
all_30 <- read.csv("../Compiled data/Correlations - 30 day rolling mean.csv") %>%
  mutate(sig = ifelse(p < alpha, T, F),
         window = "30-day rolling mean")
all_50 <- read.csv("../Compiled data/Correlations - 50 day rolling mean.csv") %>%
  mutate(sig = ifelse(p < alpha, T, F),
         window = "50-day rolling mean")
all_10 <- read.csv("../Compiled data/Correlations - 10 day rolling mean.csv") %>%
  mutate(sig = ifelse(p < alpha, T, F),
         window = "10-day rolling mean")

all <- all_30 %>%
  full_join(all_50) %>%
  full_join(all_10)

memory <- all %>% 
  filter(var %in% c("Summer epi. temperature",
                    "Summer epi. DO sat",
                    "Summer hypo. temperature",
                    "Summer VHOD",
                    "Summer hypo. DO")) %>%
  mutate(var = factor(var, levels = c("Summer epi. temperature",
                                      "Summer epi. DO sat",
                                      "Summer hypo. temperature",
                                      "Summer hypo. DO",
                                      "Summer VHOD"),
                      labels = c("Summer\nepi. temperature",
                                 "Summer\nepi. DO sat",
                                 "Summer\nhypo. temperature",
                                 "Summer\nhypo. DO",
                                 "Summer\nVHOD")),
         season = ifelse(doy >= sp[1] & doy <= sp[2], "spring", 
                         ifelse(doy >= su[1] & doy <= su[2], "summer", NA))) %>%
  group_by(var, window) %>%
  mutate(p = wilcox.test(monthly_correlation)$p.value) %>%
  filter(!is.na(season)) %>%
  group_by(LakeID, var, season, p, window) %>%
  summarize(max = max(abs(monthly_correlation))) %>%
  pivot_wider(names_from = season, values_from = max) %>%
  mutate(M = (summer - spring)) 

jpeg("../Figures/Monthly correlations - sum long concept v2 many windows.jpg", 
     width = 7, height = 3, res = 600, units = "in")
memory %>%
  filter(!var == "Summer\nepi. DO sat") %>%
  group_by(var, window) %>%
  mutate(n = length(unique(LakeID)),
         var = paste0(var, " (n = ",n,")"),
         var = factor(var, 
                      levels = c("Summer\nepi. temperature (n = 438)",
                                 "Summer\nhypo. temperature (n = 399)",
                                 "Summer\nVHOD (n = 194)",
                                 "Summer\nhypo. DO (n = 322)"),
                      labels = c("Surface-water\ntemperature (n = 438)",
                                 "Bottom-water\ntemperature (n = 399)",
                                 "Bottom-water oxygen\ndemand  (n = 194)",
                                 "Bottom-water dissolved\noxygen (n = 322)")))%>%
  group_by(var, window) %>%
  mutate(color = ifelse(p > alpha, "Insignificant", 
                        ifelse(median(M) > 0, "median > 0", "median < 0")),
         color = factor(color, levels = c("median < 0", 
                                          "Insignificant", 
                                          "median > 0"))) %>%
  ggplot(aes(y = var, x = M, color = color))+
  geom_vline(xintercept = 0, color = "grey70")+
  geom_boxplot()+
  annotate("text", x = -0.7, y = 5.8, label = "More correlated\nwith spring", 
           lineheight = .9, size = 3, hjust = 0, color = "#25469a")+
  annotate("text", x = 0.7, y = 5.8, label = "More correlated\nwith summer", 
           lineheight = .9, size = 3, hjust = 1, color = "#c16586")+
  theme_bw() +
  ggtitle("\n")+
  scale_y_discrete(limits=rev, position = "right") +
  xlab("Relative correlation with summer\nvs. spring air temperature")+
  guides(color = guide_legend(nrow = 2))+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(size= 8),
        legend.position = "none",
        legend.text = element_text(size = 8),
        legend.key.size = unit(1, "line"),
        legend.title = element_text(size = 8),
        legend.margin=margin(c(0,0,0,0)),
        legend.box.margin=margin(t = -2,l = 0, r = 0, b = -5),
        #axis.text.y=element_text(colour=c("#25469a","#25469a","#25469a","#c16586"))
        )+
  scale_color_manual(values = c("#25469a","#c16586"), 
                     name = paste0("Boxplot color\n(p < ", alpha, ")"))+
  coord_cartesian(xlim = c(-0.65,0.65), ylim = c(1,4), clip = "off")+
  facet_wrap(~window)
dev.off()
```

Step 7: Plot multiple summer definitions (SI)

```{r}
all_ja <- read.csv("../Compiled data/Correlations - 30 day rolling mean.csv") %>%
  mutate(sig = ifelse(p < alpha, T, F),
         window = "July 1 - August 31")
all_jja <- read.csv("../Compiled data/Correlations - 30 day rolling mean - JJA.csv") %>%
  mutate(sig = ifelse(p < alpha, T, F),
         window = "June 1 - August 31")
all_15 <- read.csv("../Compiled data/Correlations - 30 day rolling mean - 15th to 31st.csv") %>%
  mutate(sig = ifelse(p < alpha, T, F),
         window = "July 15 - August 31")

all <- all_ja %>%
  full_join(all_jja) %>%
  full_join(all_15)

memory <- all %>% 
  filter(var %in% c("Summer epi. temperature",
                    "Summer hypo. temperature",
                    "Summer hypo. DO")) %>%
  mutate(var = factor(var, levels = c("Summer epi. temperature",
                                      "Summer hypo. temperature",
                                      "Summer hypo. DO"),
                      labels = c("Surface-water\ntemperature",
                                 "Bottom-water\ntemperature",
                                 "Bottom-water dissolved\noxygen")),
         season = ifelse(doy >= sp[1] & doy <= sp[2], "spring", 
                         ifelse(doy >= su[1] & doy <= su[2], "summer", NA))) %>%
  group_by(var, window) %>%
  mutate(p = wilcox.test(monthly_correlation)$p.value) %>%
  filter(!is.na(season)) %>%
  group_by(LakeID, var, season, p, window) %>%
  summarize(max = max(abs(monthly_correlation))) %>%
  pivot_wider(names_from = season, values_from = max) %>%
  mutate(M = (summer - spring)) 

jpeg("../Figures/Monthly correlations - sum long concept v2 many summers.jpg", 
     width = 7, height = 3, res = 600, units = "in")
memory %>%
  group_by(var, window) %>%
  mutate(color = ifelse(p > alpha, "Insignificant", 
                        ifelse(median(M) > 0, "median > 0", "median < 0")),
         color = factor(color, levels = c("median < 0", 
                                          "Insignificant", 
                                          "median > 0"))) %>%
  ggplot(aes(y = var, x = M, color = color))+
  geom_vline(xintercept = 0, color = "grey70")+
  geom_boxplot()+
  annotate("text", x = -0.7, y = 4.4, label = "More correlated\nwith spring", 
           lineheight = .9, size = 3, hjust = 0, color = "#25469a")+
  annotate("text", x = 0.7, y = 4.4, label = "More correlated\nwith summer", 
           lineheight = .9, size = 3, hjust = 1, color = "#c16586")+
  theme_bw() +
  ggtitle("\n")+
  scale_y_discrete(limits = rev, position = "right") +
  xlab("Relative correlation with summer\nvs. spring air temperature")+
  guides(color = guide_legend(nrow = 2))+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(size = 8),
        legend.position = "none",
        legend.text = element_text(size = 8),
        legend.key.size = unit(1, "line"),
        legend.title = element_text(size = 8),
        legend.margin=margin(c(0,0,0,0)),
        legend.box.margin=margin(t = -2, l = 0, r = 0, b = -5),
        #axis.text.y=element_text(colour=c("#25469a","#25469a","#25469a","#c16586"))
        )+
  scale_color_manual(values = c("#25469a","#c16586"), 
                     name = paste0("Boxplot color\n(p < ", alpha, ")"))+
  coord_cartesian(xlim = c(-0.65,0.65), ylim = c(1, 3), clip = "off")+
  facet_wrap(~window)
dev.off()
```

Step 7: Plot partial and not-partial correlations

```{r}
all_partial <- read.csv("../Compiled data/Correlations - 30 day rolling mean.csv") %>%
  mutate(sig = ifelse(p < alpha, T, F),
         window = "Partial")
all_not_partial <- read.csv("../Compiled data/Correlations - 30 day rolling mean - not partial.csv") %>%
  mutate(sig = ifelse(p < alpha, T, F),
         window = "Not-partial")

all <- all_partial %>%
  full_join(all_not_partial)

memory <- all %>% 
  filter(var %in% c("Summer epi. temperature",
                    "Summer hypo. temperature",
                    "Summer VHOD",
                    "Summer hypo. DO")) %>%
  mutate(var = factor(var, levels = c("Summer epi. temperature",
                                      "Summer hypo. temperature",
                                      "Summer VHOD",
                                      "Summer hypo. DO"),
                      labels = c("Surface-water\ntemperature",
                                 "Bottom-water\ntemperature",
                                 "Bottom-water\noxygen demand",
                                 "Bottom-water\ndissolved oxygen")),
         season = ifelse(doy >= sp[1] & doy <= sp[2], "spring", 
                         ifelse(doy >= su[1] & doy <= su[2], "summer", NA))) %>%
  group_by(var, window) %>%
  mutate(p = wilcox.test(monthly_correlation)$p.value) %>%
  filter(!is.na(season)) %>%
  group_by(LakeID, var, season, p, window) %>%
  summarize(max = max(abs(monthly_correlation))) %>%
  pivot_wider(names_from = season, values_from = max) %>%
  mutate(M = (summer - spring)) 

jpeg("../Figures/Monthly correlations - sum long concept v2 partial not partial.jpg", 
     width = 5, height = 3, res = 600, units = "in")
memory %>%
  group_by(var, window) %>%
  mutate(color = ifelse(p > alpha, "Insignificant", 
                        ifelse(median(M) > 0, "median > 0", "median < 0")),
         color = factor(color, levels = c("median < 0", 
                                          "Insignificant", 
                                          "median > 0"))) %>%
  ggplot(aes(y = var, x = M, color = color))+
  geom_vline(xintercept = 0, color = "grey70")+
  geom_boxplot()+
  annotate("text", x = -0.7, y = 5.8, label = "More correlated\nwith spring", 
           lineheight = .9, size = 3, hjust = 0, color = "#25469a")+
  annotate("text", x = 0.7, y = 5.8, label = "More correlated\nwith summer", 
           lineheight = .9, size = 3, hjust = 1, color = "#c16586")+
  theme_bw() +
  ggtitle("\n")+
  scale_y_discrete(limits = rev, position = "right") +
  xlab("Relative correlation with summer\nvs. spring air temperature")+
  guides(color = guide_legend(nrow = 2))+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(size = 8),
        legend.position = "none",
        legend.text = element_text(size = 8),
        legend.key.size = unit(1, "line"),
        legend.title = element_text(size = 8),
        legend.margin=margin(c(0,0,0,0)),
        legend.box.margin=margin(t = -2, l = 0, r = 0, b = -5),
        #axis.text.y=element_text(colour=c("#25469a","#25469a","#25469a","#c16586"))
        )+
  scale_color_manual(values = c("#25469a","#c16586"), 
                     name = paste0("Boxplot color\n(p < ", alpha, ")"))+
  coord_cartesian(xlim = c(-0.65,0.65), ylim = c(1, 4), clip = "off")+
  facet_wrap(~window)
dev.off()

view <- all %>%
  filter(var == "Summer hypo. DO",
         LakeID == "1",
         doy == 1)

view
```

Create a plot of 1:1 correlations for SI

```{r}
medians <- memory %>%
  group_by(var) %>%
  mutate(n = length(unique(LakeID)),
         var = paste0(var, " (n = ",n,")")) %>%
  mutate(var = factor(var, 
                      levels = c("Summer\nepi. temperature (n = 438)",
                                 "Summer\nhypo. temperature (n = 399)",
                                 "Summer\nVHOD (n = 194)",
                                 "Summer\nhypo. DO (n = 322)"),
                      labels = c("Surface-water\ntemperature (n = 438)",
                                 "Bottom-water\ntemperature (n = 399)",
                                 "Bottom-water oxygen\ndemand  (n = 194)",
                                 "Bottom-water dissolved\noxygen (n = 322)")),
         layer = ifelse(grepl("Surf", var), "Surface-water", "Bottom-water")) %>%
  group_by(var, layer) %>%
  summarize(spring = median(spring, na.rm = T),
            summer = median(summer, na.rm = T),
            color = summer > spring) 

ns <- memory %>%
  group_by(var) %>%
  mutate(n = length(unique(LakeID)),
         var = paste0(var, " (n = ",n,")")) %>%
  mutate(var = factor(var, 
                      levels = c("Summer\nepi. temperature (n = 438)",
                                 "Summer\nhypo. temperature (n = 399)",
                                 "Summer\nVHOD (n = 194)",
                                 "Summer\nhypo. DO (n = 322)"),
                      labels = c("Surface-water\ntemperature (n = 438)",
                                 "Bottom-water\ntemperature (n = 399)",
                                 "Bottom-water oxygen\ndemand  (n = 194)",
                                 "Bottom-water dissolved\noxygen (n = 322)")),
         layer = ifelse(grepl("Surf", var), "Surface-water", "Bottom-water"),
         color = summer > spring) %>%
  group_by(var, color) %>%
  summarize(n = n(), 
            text = paste0("n = ", n)) %>%
  mutate(y = ifelse(color, .99, 0),
         x = ifelse(color, 0, .25),
         text = ifelse(color, paste0("Summer-driven, ", text), paste0("Spring-driven, ", text)))

one_to_one <- memory %>%
  group_by(var) %>%
  mutate(n = length(unique(LakeID)),
         var = paste0(var, " (n = ",n,")")) %>%
  mutate(var = factor(var, 
                      levels = c("Summer\nepi. temperature (n = 438)",
                                 "Summer\nhypo. temperature (n = 399)",
                                 "Summer\nVHOD (n = 194)",
                                 "Summer\nhypo. DO (n = 322)"),
                      labels = c("Surface-water\ntemperature (n = 438)",
                                 "Bottom-water\ntemperature (n = 399)",
                                 "Bottom-water oxygen\ndemand  (n = 194)",
                                 "Bottom-water dissolved\noxygen (n = 322)")),
         layer = ifelse(grepl("Surf", var), "Surface-water", "Bottom-water"),
         color = summer > spring)%>%
  ggplot(aes(x = spring, y = summer, color = color)) +
  geom_abline(slope = 1, color = "grey50") +
  geom_point() +
  facet_wrap(~var) +
  scale_color_manual(values = c("#25469a", "#c16586"))+
  scale_fill_manual(values = c("#25469a", "#c16586"))+
  xlab("Correlation with spring air temperature") +
  ylab("Correlation with summer air temeprature") +
  theme_bw() +
  ylim(0,1.05)+
  theme(legend.position = "none") +
  geom_point(data = medians, color = "white", fill = "black", size = 7, shape = 21, stroke = 0.5) +
  geom_point(data = medians, aes(fill = color), color = "white", size = 4, shape = 21, stroke = 1) + 
  geom_text(aes(x = x, y = y, label = text), data = ns, hjust = 0, vjust = 0) 

jpeg("../Figures/one_to_one.jpg", width = 5, height = 5, units = "in", res = 300)
one_to_one
dev.off()
```

