---
title: "Air temp analysis"
author: "Abby Lewis"
date: "2023-05-22"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

This file calculates "seasonal ecological memory" as the maximum absolute correlation between our summer focal variables and 30-day rolling mean air temperature during July--August (summer) minus the maximum correlation during spring (March-May).

Table of contents:

-   Step 1: Load packages and data
-   Step 2: Calculate seasonal ecological memory
-   Step 3: Generate Figure 1 (summary boxplot)
-   Step 4: Plot rolling means (hypolimnetic variables)
-   Step 5: Identify most influential dates
-   Step 6: Plot multiple rolling-mean windows (SI)

Step 1: Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)

#Set global options
sp <- yday(c("2022-03-01", "2022-05-31"))
su <- yday(c("2022-07-01", "2022-08-31"))
alpha <- 0.00001

#Load saved data
lat_long <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/1530/1/fadd3eaa25b5fdd1fc4efba70e660579")
all <- read.csv("../Compiled data/Correlations - 30 day rolling mean - surf bot.csv") %>%
  mutate(sig = ifelse(p < alpha, T, F))
onset <- read.csv("../Compiled data/stratification_onset.csv")

# Make a directory for Figures if there isn't one
if (!dir.exists("../Figures")){dir.create("../Figures")}

#Labs for plots
levels_no_n = c("Summer epi. temperature", 
                "Summer hypo. temperature",
                "Summer VHOD",
                "Summer hypo. DO")

labels_no_n = c("Surface-water\ntemperature",
                "Bottom-water\ntemperature",
                "Bottom-water\noxygen demand",
                "Bottom-water\ndissolved oxygen")

stratified <- read_csv("../Compiled data/stratified_lakes.csv")
```

Step 2: Calculate seasonal ecological memory

```{r}
memory <- all %>% 
  filter(var %in% levels_no_n) %>%
  mutate(var = factor(var, 
                      levels = levels_no_n,
                      labels = labels_no_n),
         season = ifelse(doy >= sp[1] & doy <= sp[2], "spring", 
                         ifelse(doy >= su[1] & doy <= su[2], "summer", NA))) %>%
  filter(!is.na(season)) %>%
  group_by(LakeID, var, season) %>%
  summarize(max = max(abs(monthly_correlation))) %>%
  pivot_wider(names_from = season, values_from = max) %>%
  mutate(M = (summer - spring)) %>%
  group_by(var) %>%
  mutate(p = wilcox.test(M)$p.value)

#memory_sat <- hypo_do_sat %>%
#  mutate(season = ifelse(doy >= sp[1] & doy <= sp[2], "spring", 
#                         ifelse(doy >= su[1] & doy <= su[2], "summer", NA))) %>%
#  filter(!is.na(season)) %>%
#  group_by(LakeID, var, season) %>%
#  summarize(max = max(abs(monthly_correlation))) %>%
#  pivot_wider(names_from = season, values_from = max) %>%
#  mutate(M = (summer - spring)) %>%
#  group_by(var) %>%
#  mutate(p = wilcox.test(M)$p.value)
```

Step 3: Generate Figure 1 (summary boxplot)

```{r}
jpeg("../Figures/Monthly correlations - sum long concept v2 surf bot.jpg", width = 3.5, height = 3, res = 600, units = "in")
memory %>%
  filter(!var == "Summer\nepi. DO sat") %>%
  group_by(var) %>%
  mutate(n = length(unique(LakeID)),
         var = paste0(var, " (n = ",n,")")) %>%
  mutate(var = factor(var, 
                      levels = c("Surface-water\ntemperature (n = 561)",
                                 "Bottom-water\ntemperature (n = 568)",
                                 "Bottom-water\noxygen demand (n = 194)",
                                 "Bottom-water\ndissolved oxygen (n = 480)"),
                      labels = c("Surface-water\ntemperature (n = 561)",
                                 "Bottom-water\ntemperature (n = 568)",
                                 "Bottom-water oxygen\ndemand (n = 194)",
                                 "Bottom-water dissolved\noxygen (n = 480)")))%>%
  group_by(var) %>%
  mutate(color = ifelse(p > alpha, "Insignificant", 
                        ifelse(median(M) > 0, "median > 0", "median < 0")),
         color = factor(color, levels = c("median < 0", 
                                          "Insignificant", 
                                          "median > 0"))) %>%
  ggplot(aes(y = var, x = M, color = color))+
  geom_vline(xintercept = 0, color = "grey70")+
  geom_boxplot()+
  annotate("text", x = -0.7, y = 5, label = "More correlated\nwith spring", 
           lineheight = .9, size = 3, hjust = 0, color = "#25469a")+
  annotate("text", x = 0.7, y = 5, label = "More correlated\nwith summer", 
           lineheight = .9, size = 3, hjust = 1, color = "#c16586")+
  theme_bw() +
  ggtitle("\n")+
  scale_y_discrete(limits=rev, position = "right") +
  xlab("Relative correlation with summer\nvs. spring air temperature")+
  guides(color = guide_legend(nrow = 2))+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(size= 8),
        legend.position = "none",
        legend.text = element_text(size = 8),
        legend.key.size = unit(1, "line"),
        legend.title = element_text(size = 8),
        legend.margin = ggplot2::margin(c(0,0,0,0)),
        legend.box.margin=ggplot2::margin(t = -2,l = 0, r = 0, b = -5),
        #axis.text.y=element_text(colour=c("#25469a","#25469a","#25469a","#c16586"))
        )+
  scale_color_manual(values = c("#25469a","#c16586"), 
                     name = paste0("Boxplot color\n(p < ", alpha, ")"))+
  coord_cartesian(xlim = c(-0.65,0.65), ylim = c(1,4), clip = "off")
dev.off()
```

Step 3b: Compare hypo DO concentration (mg/L) to hypo DO saturation (%)

```{r}
#jpeg("../Figures/Monthly correlations - DO sat comp surf bot.jpg", width = 3.5, height = 2, res = 600, units = "in")
#memory_sat %>%
#  mutate(var = "Summer hypo. DO\nsaturation") %>%
#  full_join(memory %>% 
#              filter(var == "Bottom-water\ndissolved oxygen") %>%
#              mutate(var = "Summer hypo. DO\nconcentration")) %>%
#  group_by(var) %>%
#  mutate(n = length(unique(LakeID)),
#         var = paste0(var, " (n = ",n,")")) %>%
#  group_by(var) %>%
#  mutate(color = ifelse(p > alpha, "Insignificant", 
#                        ifelse(median(M) > 0, "median > 0", "median < 0")),
#         color = factor(color, levels = c("median < 0", 
#                                          "Insignificant", 
#                                          "median > 0"))) %>%
#  ggplot(aes(y = var, x = M, color = color))+
#  geom_vline(xintercept = 0, color = "grey70")+
#  geom_boxplot()+
#  annotate("text", x = -0.7, y = 3, label = "More correlated\nwith spring", 
#           lineheight = .9, size = 3, hjust = 0, color = "#25469a")+
#  annotate("text", x = 0.7, y = 3, label = "More correlated\nwith summer", 
#           lineheight = .9, size = 3, hjust = 1, color = "#c16586")+
#  theme_bw() +
#  ggtitle("\n")+
#  scale_y_discrete(limits=rev, position = "right") +
#  xlab("Relative correlation with summer\nvs. spring air temperature")+
#  guides(color = guide_legend(nrow = 2))+
#  theme(axis.title.y = element_blank(),
#        axis.title.x = element_text(size= 8),
#        legend.position = "none",
#        legend.text = element_text(size = 8),
#        legend.key.size = unit(1, "line"),
#        legend.title = element_text(size = 8),
#        legend.margin=ggplot2::margin(c(0,0,0,0)),
#        legend.box.margin=ggplot2::margin(t = -2,l = 0, r = 0, b = -5),
#        #axis.text.y=element_text(colour=c("#25469a","#25469a","#25469a","#c16586"))
#        )+
#  scale_color_manual(values = c("#25469a","#c16586"), 
#                     name = paste0("Boxplot color\n(p < ", alpha, ")"))+
#  coord_cartesian(xlim = c(-0.65,0.65), ylim = c(1,2), clip = "off")
#dev.off()
```

Step 4: Plot rolling means (hypolimnetic variables)

```{r}
#Function to deliniate periods of significant vs insignificant correlations
group_fun <- function(sig) {
  group <- rep(1, length(sig))
  for (i in 2:length(sig)) {
    if(sig[i] == sig[i - 1]) {
      group[i] <- group[i - 1]
    } else {
      group[i] <- group[i - 1] + 1
    }
  }
  return(group)
}

#Calculate mean and SD
all_plot <- all %>%
  mutate(strat = LakeID %in% stratified$LakeID) %>%
  group_by(var, strat) %>%
  mutate(n = length(unique(LakeID)), #Some lakes are missing from January because our air temp only goes back to 1980
         var = paste0(var, " (n = ",n,")")) %>%
  group_by(var, doy, strat) %>%
    mutate(p = wilcox.test(monthly_correlation)$p.value,
           sig = p < 0.001) %>%
  group_by(var, doy, sig, strat) %>%
  summarise(mean = mean(monthly_correlation, na.rm = T),
            sd = sd(monthly_correlation, na.rm = T)) 

#Use group function
all_plot$group <- 1
for(i in 1:length(unique(all_plot$var))){
  all_plot$group[all_plot$var == unique(all_plot$var)[i]] <- 
    group_fun(all_plot$sig[all_plot$var == unique(all_plot$var)[i]])
}

#To get smooth ribbons, we need to manually duplicate datapoints
all_plot1 <- all_plot %>%
  mutate(DateTime = as.Date("2022-01-01") + days(doy))
all_plot2 <- all_plot %>%
  ungroup() %>%
  mutate(DateTime = as.Date("2022-01-01") + days(doy) + hours(23) + minutes(59),
         mean = ifelse(!is.na(lead(var)) & lead(var) == var, lead(mean), mean),
         sd = ifelse(lead(var) == var, lead(sd), sd))

#Plot hypolimnetic variables
hypo_vars_strat <- all_plot1 %>%
  full_join(all_plot2) %>%
  filter(doy>=30) %>% #remove first 29 days (these include previous year) 
  filter(strat == T) %>%
  filter(var %in% c("Summer VHOD (n = 192)",
                    #"Summer epi. temperature (n = 469)",
                    "Summer hypo. DO (n = 383)",
                    "Summer hypo. temperature (n = 468)")) %>%
  mutate(var = factor(var, 
                      levels = c(#"Summer epi. temperature (n = 469)",
                                 "Summer hypo. temperature (n = 468)",
                                 "Summer VHOD (n = 192)",
                                 "Summer hypo. DO (n = 383)"
                                 ),
                      labels = c(#"Summer surface-water\ntemperature (n = 469)",
                                 "Summer bottom-water\ntemperature (n = 468)",
                                 "Summer bottom-water\noxygen demand (n = 192)",
                                 "Summer bottom-water\nDO (n = 383)")),
         layer = ifelse(grepl("epi", var), "epi", "hypo")) %>%
  filter(!var == "Summer bottom-water\noxygen demand") %>%
  ggplot() + 
  geom_hline(yintercept=0, color = "grey60", linewidth = 0.3)+
  geom_vline(xintercept = as_datetime("2022-01-01")+
               days(median(onset$Onset, na.rm = T)),
             color = "grey50", lty = "dashed")+
  geom_ribbon(aes(x = DateTime, 
                  ymin = mean - sd, ymax = mean + sd,
                  fill = sig, group = group),
              alpha = 0.5) +
  geom_line(aes(x = DateTime, y = mean)) +
  xlab("Air temperature date (right-aligned 30-day rolling mean)")+
  ylab("Correlation")+
  theme_bw()+
  scale_fill_manual(values = c("grey40","#0FA9E6")) +
  labs(fill = paste0("p < ",alpha)) +
  facet_wrap(~var, nrow = 1, drop = F) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        strip.background = element_blank()) +
  ggtitle("Dimictic and warm-monomictic lakes")

hypo_vars_unstrat <- all_plot1 %>%
  full_join(all_plot2) %>%
  filter(doy>=30) %>% #remove first 29 days (these include previous year)
  filter(var %in% c(#"Summer epi. temperature (n = 92)",
                    "Summer hypo. DO (n = 97)",
                    "Summer hypo. temperature (n = 92)")) %>%
  mutate(var = factor(var, 
                      levels = c(#"Summer epi. temperature (n = 92)",
                                 "Summer hypo. temperature (n = 92)",
                                 "Summer VHOD (n = 2)",
                                 "Summer hypo. DO (n = 97)"
                                 ),
                      labels = c(#"Summer surface-water\ntemperature (n = 92)",
                                 "Summer bottom-water\ntemperature (n = 92)",
                                 "Summer bottom-water\noxygen demand",
                                 "Summer bottom-water\nDO (n = 97)")),
         layer = ifelse(grepl("epi", var), "epi", "hypo")) %>%
  filter(!var == "Summer bottom-water\noxygen demand") %>%
  ggplot() + 
  geom_hline(yintercept=0, color = "grey60", linewidth = 0.3)+
  geom_ribbon(aes(x = DateTime, 
                  ymin = mean - sd, ymax = mean + sd,
                  fill = sig, group = group),
              alpha = 0.5) +
  geom_line(aes(x = DateTime, y = mean)) +
  xlab("Air temperature date (right-aligned 30-day rolling mean)")+
  ylab("Correlation")+
  theme_bw()+
  scale_fill_manual(values = c("grey40","#0FA9E6"))+
  labs(fill = paste0("p < ",alpha)) +
  facet_wrap(~var, nrow = 1, drop = F) +
  theme(legend.position = "bottom",
        legend.margin=ggplot2::margin(c(0,0,0,0)),
        strip.background = element_blank())+
  ggtitle("Polymicitc lakes")

#Save plot
jpeg("../Figures/Monthly correlations - hypo daily roll mean sd surf bot.jpg", 
     width = 7, height = 4, res = 300, units = "in")
ggarrange(hypo_vars_strat, hypo_vars_unstrat, nrow = 2, align = "hv", common.legend = T, legend = "bottom")
dev.off()
```
Export dates

```{r}
dates_to_plot <- all %>%
  mutate(season = ifelse(doy >= sp[1] & doy <= sp[2], "spring", 
                         ifelse(doy >= su[1] & doy <= su[2], "summer", NA))) %>%
  filter(!is.na(season)) %>%
  group_by(var, LakeID, season) %>%
  summarize(max_doy = doy[which.max(abs(monthly_correlation))]) 

write.csv(dates_to_plot, 
          "../Compiled data/strongest_correlation_dates_all_met_3m.csv", 
          row.names = F)
```

