---
title: "Air temp analysis"
author: "Abby Lewis"
date: "2023-05-22"
output: html_document
---

This file performs a driver analysis to characterize the extent to which lake morphometric factos explain variation in the strength of seasonal ecological memory

Table of contents: 
- Step 1: Load packages and data 
- Step 2: Calculate seasonal ecological memory
- Step 3: Run driver analysis
- Step 4: Plot identified drivers

Step 1: Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ggpubr)
source("lmer_functions.R")

#Set global options
sp <- yday(c("2022-03-01", "2022-05-31"))
su <- yday(c("2022-07-01", "2022-08-31"))

#Load saved data
with_temp = read.csv("../Compiled data/All_data_annual.csv")
lat_long <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/1530/1/fadd3eaa25b5fdd1fc4efba70e660579")
all <- read.csv("../Compiled data/Correlations - 30 day rolling mean.csv") 
#Buoyancy frequency
bf = read.csv("../Compiled data/summer_averages_wi.csv")%>%
  group_by(LakeID) %>%
  summarize(buoyancy_freq = median(buoyancy_freq, na.rm = T))

# Make a directory for Figures if there isn't one
if (!dir.exists("../Figures")){dir.create("../Figures")}
```

Step 2: Calculate seasonal ecological memory

```{r}
memory <- all %>% 
  filter(var %in% c("Summer epi. temperature",
                    "Summer hypo. temperature",
                    "Summer VHOD",
                    "Summer hypo. DO")) %>%
  mutate(var = factor(var, levels = c("Summer epi. temperature",
                                      "Summer hypo. temperature",
                                      "Summer VHOD",
                                      "Summer hypo. DO"),
                      labels = c("Summer epi. temperature",
                                 "Summer bottom-water\ntemperature",
                                 "Summer VHOD",
                                 "Summer bottom-water\ndissolved oxygen")),
         season = ifelse(doy >= sp[1] & doy <= sp[2], "spring", 
                         ifelse(doy >= su[1] & doy <= su[2], "summer", NA))) %>%
  group_by(var) %>%
  mutate(p = wilcox.test(monthly_correlation)$p.value) %>%
  filter(!is.na(season)) %>%
  group_by(LakeID, var, season, p) %>%
  summarize(max = max(abs(monthly_correlation))) %>%
  pivot_wider(names_from = season, values_from = max) %>%
  mutate(M = (summer - spring)) 
```

Step 3: Run driver analysis

```{r}
mem_lm <- memory %>%
  select(LakeID, var, M) %>%
  left_join(bf) %>%
  left_join(lat_long) %>%
  mutate(log_SA = log(SurfaceArea_ha),
         log_max_depth = log(MaximumDepth_m)) %>%
  pivot_wider(names_from = var, values_from = M) %>%
  rename(epi_temp = `Summer epi. temperature`,
         hypo_temp = `Summer bottom-water\ntemperature`,
         hypo_do = `Summer bottom-water\ndissolved oxygen`,
         vhod = `Summer VHOD`)

responses <- c("epi_temp")
potential_drivers <- c("log_SA", "log_max_depth", "buoyancy_freq")
mem_lm_df <- standardize_data(mem_lm, responses, potential_drivers)
fm1 <- lm(epi_temp ~ log_SA*log_max_depth*buoyancy_freq, 
          data = mem_lm_df,
          na.action = "na.fail")
dd <- MuMIn::dredge(fm1)
subset(dd, delta < 2)
export_epi <- subset(dd, delta < 2) %>% 
  mutate(var = "Surface-water temperature") %>%
  filter(!is.na(`(Intercept)`))
mod_lm_epi_temp <- lm(epi_temp ~ log_SA + log_max_depth +
                    log_max_depth:log_SA, 
             data = mem_lm_df) 
summary(mod_lm_epi_temp)
#No significant main effect relationships

responses <- c("hypo_temp")
mem_lm_df <- standardize_data(mem_lm,responses,potential_drivers)
fm1 <- lm(hypo_temp ~ log_SA * log_max_depth * buoyancy_freq, 
          data = mem_lm_df,
          na.action = "na.fail")
dd <- MuMIn::dredge(fm1)
subset(dd, delta < 2)
export_hypo <- subset(dd, delta < 2) %>% 
  mutate(var = "Bottom-water temperature") %>%
  filter(!is.na(`(Intercept)`))
mod_lm_temp <- lm(hypo_temp ~ log_SA + buoyancy_freq + log_max_depth +
                    log_max_depth:log_SA, 
             data = mem_lm_df) 
summary(mod_lm_temp)

responses <- c("hypo_do")
mem_lm_df <- standardize_data(mem_lm,responses,potential_drivers)
fm1 <- lm(hypo_do ~ log_SA * log_max_depth * buoyancy_freq, 
          data = mem_lm_df,
          na.action = "na.fail")
dd <- MuMIn::dredge(fm1)
subset(dd, delta < 2)
export_do <- subset(dd, delta < 2) %>% 
  mutate(var = "Bottom-water dissolved oxygen") %>%
  filter(!is.na(`(Intercept)`))
mod_lm_do <- lm(hypo_do ~ log_SA + buoyancy_freq + log_max_depth + log_SA:buoyancy_freq, data = mem_lm_df) 
summary(mod_lm_do)

responses <- c("vhod")
mem_lm_df <- standardize_data(mem_lm, responses, potential_drivers)
fm1 <- lm(vhod ~ log_SA * log_max_depth * buoyancy_freq, 
          data = mem_lm_df,
          na.action = "na.fail")
dd <- MuMIn::dredge(fm1)
subset(dd, delta < 2)
export_vhod <- subset(dd, delta < 2) %>% 
  mutate(var = "Bottom-water oxygen demand") %>%
  filter(!is.na(`(Intercept)`))
#Intercept only model is best

export_all <- export_epi %>%
  full_join(export_hypo) %>%
  full_join(export_do) %>%
  full_join(export_vhod)
export <- as.data.frame(export_all) %>%
  relocate(var) %>%
  select(-weight, -delta) %>%
  select(where(~sum(!is.na(.x)) > 0)) %>%
  mutate(across(is.numeric, round, 3),
         across(c(logLik, AICc), round, 1),
         across(everything(), ~ifelse(is.na(.),"",.))) %>%
  rename(`Buoyancy freq.` = buoyancy_freq,
         `Log-likelihood` = logLik,
         `Intercept` = `(Intercept)`,
         `Max. depth` = log_max_depth,
         `Surface area` = `log_SA`,
         `Buoyancy freq. :\nmax. depth` = `buoyancy_freq:log_max_depth`,
         `Buoyancy freq. :\nsurface area` = `buoyancy_freq:log_SA`,
         `Max. depth :\nsurface area` = `log_max_depth:log_SA`)
write.csv(export, "../Compiled data/driver analysis results.csv", row.names = F)
```

Step 4: Plot identified drivers

```{r}
mem_plot <- memory %>%
  filter(var %in% c("Summer bottom-water\ntemperature",
                    "Summer bottom-water\ndissolved oxygen")) %>%
  left_join(bf) %>%
  left_join(lat_long) %>%
  mutate(SA_class = ifelse(SurfaceArea_ha > 100, "> 100 ha", "≤ 100 ha")) %>%
  filter(!is.na(SurfaceArea_ha)) %>%
  ggplot(aes(x = buoyancy_freq, y = M))+
  geom_hline(yintercept = 0)+
  geom_point(aes(color = SA_class, shape = SA_class))+
  geom_smooth(method = "lm", aes(color = SA_class))+
  theme_bw() +
  xlab("Maximum buoyancy frequency (1/s)")+
  scale_color_manual(name = "Lake size", values = c("#202C39", "#65A47D"))+
  scale_shape_discrete(name = "Lake size")+
  facet_wrap(~var)+
  ylab("\n\n\n")+
  geom_text(x = -0.0005, y = 0.65, label = "More correlated\nwith summer", 
           hjust = 1, lineheight = .9, size = 3, 
           data = data.frame(var = factor("Summer bottom-water\ntemperature",
                                          levels = c("Summer bottom-water\ntemperature",
                                                     "Summer bottom-water\ndissolved oxygen"))))+
  geom_text(x = -0.0005, y = -0.65, label = "More correlated\nwith spring", 
           hjust = 1, lineheight = .9, size = 3, 
           data = data.frame(var = factor("Summer bottom-water\ntemperature",
                                          levels = c("Summer bottom-water\ntemperature",
                                                     "Summer bottom-water\ndissolved oxygen"))))+
  coord_cartesian(ylim = c(-0.65,0.65), clip = "off")

jpeg("../Figures/Memory drivers.jpg", res = 300, width = 5, height = 3, units = "in")
mem_plot + 
  guides(shape = guide_legend(override.aes = list(size = 3)))+
  theme(legend.position = "bottom",
        legend.margin=margin(c(0,0,0,0)))
dev.off()
```

