---
title: "Air temp trends"
author: "Abby Lewis"
date: "2023-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(trend)
library(ggh4x)
library(ggpubr)
library(randomForest)

# Make a directory for Figures if there isn't one
if (!dir.exists("../Figures")){dir.create("../Figures")}

# Load data
lat_long = read.csv("https://pasta-s.lternet.edu/package/data/eml/edi/1029/9/fadd3eaa25b5fdd1fc4efba70e660579")
met_by_bot_temp <- read_csv("../Compiled data/climate_era5_roll_met_daily - summer_bot_temp_years.csv")
met_by_surf_temp <- read_csv("../Compiled data/climate_era5_roll_met_daily - summer_surf_temp_years.csv")
met_by_bot_do <- read_csv("../Compiled data/climate_era5_roll_met_daily - summer_bot_DO_years.csv")
max_cor_dates <- read.csv("../Compiled data/strongest_correlation_dates_all_met_3m.csv")
mem <- read_csv("../Compiled data/memory_3m.csv") %>%
  mutate(var = recode(var, 
                      "Summer bottom-water\ntemperature" = "summer_bot_mean",
                      "Summer bottom-water\ndissolved oxygen" = "summer_bot_do",
                      "Summer epi. temperature" = "summer_surf_mean"))
summer_bot_mean <- read.csv("../Compiled data/summer_bot_temp_mean_trends_3m.csv") %>%
  rename(summer_bot_mean = trend) %>%
  dplyr::select(summer_bot_mean, LakeID)
summer_surf_mean <- read.csv("../Compiled data/summer_surf_temp_trends_3m.csv") %>%
  rename(summer_surf_mean = trend) %>%
  dplyr::select(summer_surf_mean, LakeID)
summer_surf_do <- read.csv("../Compiled data/summer_surf_DO_trends_3m.csv") %>%
  rename(summer_surf_do = trend) %>%
  dplyr::select(summer_surf_do, LakeID)
summer_bot_do <- read.csv("../Compiled data/summer_bot_DO_trends_3m.csv") %>%
  rename(summer_bot_do = trend) %>%
  dplyr::select(summer_bot_do, LakeID)
summer_surf_tp <- read.csv("../Compiled data/summer_surf_tp_trends_3m.csv") %>%
  rename(summer_surf_tp = trend) %>%
  dplyr::select(summer_surf_tp, LakeID)
summer_dif <- read.csv("../Compiled data/summer_dif_trends_3m.csv") %>%
  rename(summer_dif = trend) %>%
  dplyr::select(summer_dif, LakeID)
summer_dif_do <- read.csv("../Compiled data/summer_dif_DO_trends_3m.csv") %>%
  rename(summer_dif = trend) %>%
  dplyr::select(summer_dif, LakeID)

met_by_bot_temp %>%
  mutate(doy = as.Date("2022-01-01") + doy - 1) %>%
  ggplot(aes(x = doy, y = trend, group = LakeID)) +
  geom_line(alpha=0.01)+
  facet_wrap(~Variable, scales = "free_y")
```

Compare

```{r}
data <- summer_bot_mean %>%
  full_join(summer_surf_mean) %>%
  full_join(summer_surf_do) %>%
  full_join(summer_bot_do) %>%
  pivot_longer(cols = c(summer_bot_mean, summer_surf_mean, summer_surf_do, summer_bot_do),
               names_to = "Variable",
               values_to = "trend") %>%
  left_join(mem %>% select(LakeID, var, M) %>% distinct(), by = c("LakeID","Variable" = "var")) %>%
  mutate(layer = ifelse(str_detect(Variable, "surf"), 
                        "Surface", "Bottom"),
         var = ifelse(str_detect(Variable, "mean"), 
                      "Temperature trend\n(ºC/decade)", "DO trend\n(mg/L/decade)"),
         var = factor(var,
                      levels = c("Temperature trend\n(ºC/decade)", "DO trend\n(mg/L/decade)"))) %>%
  group_by(var, LakeID) %>%
  mutate(M = M[layer == "Bottom"]) %>%
  select(-Variable) %>%
  pivot_wider(names_from = layer,
              values_from = trend)

cors <- data %>%
  group_by(var) %>%
  summarize(cor = cor(Surface, Bottom, method = "spearman", use = "complete.obs"))

min_do <- min(c(data$Bottom[data$var == "DO trend\n(mg/L/decade)"], 
                data$Surface[data$var == "DO trend\n(mg/L/decade)"]),
              na.rm = T)
max_do <- max(c(data$Bottom[data$var == "DO trend\n(mg/L/decade)"], 
                data$Surface[data$var == "DO trend\n(mg/L/decade)"]),
              na.rm = T)
min_temp <- min(c(data$Bottom[data$var == "Temperature trend\n(ºC/decade)"], 
                  data$Surface[data$var == "Temperature trend\n(ºC/decade)"]),
                na.rm = T)
max_temp <- max(c(data$Bottom[data$var == "Temperature trend\n(ºC/decade)"], 
                  data$Surface[data$var == "Temperature trend\n(ºC/decade)"]),
                na.rm = T)
mean_do <- mean(c(data$Bottom[data$var == "DO trend\n(mg/L/decade)"], 
                data$Surface[data$var == "DO trend\n(mg/L/decade)"]),
              na.rm = T)
sd_do <- sd(c(data$Bottom[data$var == "DO trend\n(mg/L/decade)"], 
                data$Surface[data$var == "DO trend\n(mg/L/decade)"]),
              na.rm = T)
mean_temp <- mean(c(data$Bottom[data$var == "Temperature trend\n(ºC/decade)"], 
                  data$Surface[data$var == "Temperature trend\n(ºC/decade)"]),
                na.rm = T)
sd_temp <- sd(c(data$Bottom[data$var == "Temperature trend\n(ºC/decade)"], 
                  data$Surface[data$var == "Temperature trend\n(ºC/decade)"]),
                na.rm = T)

jpeg("../Figures/spatial_mismatch.jpg", width = 6, height = 3, units = "in", res = 300)
data %>%
  ggplot(aes(x = Bottom, y = Surface)) +
  geom_abline()+
  geom_point(alpha = 0.5) +
  geom_label(data = cors, aes(x = Inf, y = Inf, label =  paste0("\u03C1 = ", round(cor,2))), 
             hjust = 1, vjust = 1)+
  facet_wrap(~var, scales = "free")+
  ggh4x::facetted_pos_scales(
    x = list(
      scale_x_continuous(limits = c(mean_do - 2*sd_do, 
                                    mean_do + 2*sd_do)),
      scale_x_continuous(limits = c(mean_temp - 2*sd_temp, 
                                    mean_temp + 2*sd_temp))
    ),
    y = list(
      scale_y_continuous(limits = c(mean_do - 2*sd_do, 
                                    mean_do + 2*sd_do)),
      scale_y_continuous(limits = c(mean_temp - 2*sd_temp, 
                                    mean_temp + 2*sd_temp))
    )
  )+
  theme_bw()
dev.off()

data %>%
  ggplot(aes(x = Bottom, y = Surface)) +
  geom_abline()+
  geom_point(alpha = 0.8, aes(color = M)) +
  geom_label(data = cors, aes(x = Inf, y = Inf, label =  paste0("\u03C1 = ", round(cor,2))), 
             hjust = 1, vjust = 1)+
  facet_wrap(~var, scales = "free")+
  theme_bw()+
  scale_color_gradientn(colors = c("#25469a", "#735690","#c16586"), 
                        limits = c(-0.7, 0.7))+
  ggh4x::facetted_pos_scales(
    x = list(
      scale_x_continuous(limits = c(mean_do - 2*sd_do, 
                                    mean_do + 2*sd_do)),
      scale_x_continuous(limits = c(mean_temp - 2*sd_temp, 
                                    mean_temp + 2*sd_temp))
    ),
    y = list(
      scale_y_continuous(limits = c(mean_do - 2*sd_do, 
                                    mean_do + 2*sd_do)),
      scale_y_continuous(limits = c(mean_temp - 2*sd_temp, 
                                    mean_temp + 2*sd_temp))
    )
  )

a <- data %>%
  filter(var == "Temperature trend\n(ºC/decade)") %>%
  ggplot(aes(x = Bottom*10, y = Surface*10)) +
  geom_abline(color = "grey70")+
  geom_hline(yintercept = 0, color = "grey80")+
  geom_vline(xintercept = 0, color = "grey80")+
  geom_point(alpha = 0.5) +
  ylab("Surface temperature trend\n(ºC/decade)")+
  xlab("Bottom temperature trend\n(ºC/decade)")+
  geom_label(data = cors %>%
               filter(var == "Temperature trend\n(ºC/decade)"), 
             aes(x = -10, y = 4, label =  paste0("\u03C1 = ", round(cor,2))), 
             hjust = 0)+
  theme_bw(base_size = 10)

b <- summer_dif %>% 
  ggplot(aes(x = summer_dif*10)) +
  geom_vline(aes(xintercept = 0), color = "grey70")+
  geom_hline(yintercept = 0, color = "grey80")+
  geom_vline(aes(xintercept = median(summer_dif*10, na.rm = T)), lty = "dotted")+
  geom_density(linewidth = 1)+
  geom_label(aes(x = -3, y = 0.35, 
                 label = paste0("Median = +", 
                                round(median, 2),
                                " ºC/decade")), 
             hjust = 0,
             data = . %>% summarize(median = median(summer_dif, na.rm = T)*10))+
  xlab("Trend in thermal difference between\n surface and bottom waters (ºC/decade)")+
  ylab("Density") +
  theme_bw(base_size = 10)

jpeg("../Figures/temp_mismatch.jpg", width = 6, height = 3, units = "in", res = 300)
ggarrange(a, b, labels = "auto")
dev.off()
```


Lm

```{r}
met_trends_matched <- max_cor_dates %>%
  filter(var == "Summer hypo. temperature") %>%
  left_join(met_by_bot_temp, 
            by = c("LakeID", "max_doy" = "doy"), 
            relationship = "many-to-many") %>%
  group_by(season, var) %>%
  rename(cor_date_var = var)

df_for_rf <- summer_bot_mean %>%
  left_join(met_trends_matched %>% 
              select(LakeID, trend, season, Variable, cor_date_var) %>%
              arrange(LakeID) %>%
              pivot_wider(names_from = Variable,
                          values_from = trend),
            by = "LakeID") %>%
  pivot_wider(names_from = season,
              values_from = Air_temp_C:Shortwave_mJm2) %>%
  left_join(mem %>% filter(var == "summer_bot_mean") %>% select(M, LakeID), by = "LakeID") %>%
  mutate(M_binary = ifelse(M > 0, "> 0", "<= 0"))

responses <- c("summer_bot_mean")
potential_drivers <- c("Air_temp_C_spring", "Air_temp_C_summer")
mem_rf_df_bot_temp <- df_for_rf %>%
  select(all_of(c(responses, potential_drivers))) %>%
  na.omit() %>%
  #mutate(M = (M + 1)/2) %>%
  mutate(target = "Bottom-water\ntemperature")
lm1 <- lm(summer_bot_mean ~ Air_temp_C_spring + Air_temp_C_summer, 
          data = mem_rf_df_bot_temp)
summary(lm1)

mem_rf_df_bot_temp %>%
  ggplot(aes(x = Air_temp_C_spring, y = summer_bot_mean)) +
  geom_point(alpha = 0.2)+
  geom_smooth(method = "gam")

# Surface
met_trends_matched <- max_cor_dates %>%
  filter(var == "Summer epi. temperature") %>%
  left_join(met_by_surf_temp, 
            by = c("LakeID", "max_doy" = "doy"), 
            relationship = "many-to-many") %>%
  group_by(season, var) %>%
  rename(cor_date_var = var)

df_for_rf <- summer_surf_mean %>%
  left_join(met_trends_matched %>% 
              select(LakeID, trend, season, Variable) %>%
              arrange(LakeID) %>%
              pivot_wider(names_from = Variable,
                          values_from = trend))  %>%
  pivot_wider(names_from = season,
              values_from = Air_temp_C:Shortwave_mJm2) %>%
  left_join(mem %>% select(M, LakeID, var) %>% filter(var == "summer_surf_mean")) %>%
  mutate(M_binary = ifelse(M > 0, "> 0", "<= 0"))

responses <- c("summer_surf_mean")
potential_drivers <- c("Air_temp_C_spring", "Air_temp_C_summer")
mem_rf_df_surf_temp <- df_for_rf %>%
  select(all_of(c(responses, potential_drivers))) %>%
  na.omit() %>%
  #mutate(M = (M + 1)/2) %>%
  mutate(target = "Surface-water\ntemperature")
lm1 <- lm(summer_surf_mean ~ Air_temp_C_spring + Air_temp_C_summer, 
          data = mem_rf_df_surf_temp)
summary(lm1)

# DO
met_trends_matched <- max_cor_dates %>%
  filter(var == "Summer hypo. DO") %>%
  left_join(met_by_bot_do, 
            by = c("LakeID", "max_doy" = "doy"), 
            relationship = "many-to-many") %>%
  group_by(season, var) %>%
  rename(cor_date_var = var)

df_for_rf <- summer_bot_do %>%
  left_join(met_trends_matched %>% 
              select(LakeID, trend, season, Variable) %>%
              arrange(LakeID) %>%
              pivot_wider(names_from = Variable,
                          values_from = trend))  %>%
  pivot_wider(names_from = season,
              values_from = Air_temp_C:Shortwave_mJm2) %>%
  left_join(mem %>% select(M, LakeID, var) %>% filter(var == "summer_bot_mean")) %>%
  mutate(M_binary = ifelse(M > 0, "> 0", "<= 0"))

responses <- c("summer_bot_do")
potential_drivers <- c("Air_temp_C_spring", "Air_temp_C_summer")
mem_rf_df_bot_do <- df_for_rf %>%
  select(all_of(c(responses, potential_drivers))) %>%
  na.omit() %>%
  #mutate(M = (M + 1)/2) %>%
  mutate(target = "Bottom-water\ndissolved oxygen")
lm1 <- lm(summer_bot_do ~ Air_temp_C_spring + Air_temp_C_summer, 
          data = mem_rf_df_bot_do)
summary(lm1)
```
RF: bottom-water temperature

```{r}
met_trends_matched <- max_cor_dates %>%
  filter(var == "Summer hypo. temperature") %>%
  left_join(met_by_bot_temp, 
            by = c("LakeID", "max_doy" = "doy"), 
            relationship = "many-to-many") %>%
  group_by(season, var) %>%
  rename(cor_date_var = var)

df_for_rf <- summer_bot_mean %>%
  left_join(met_trends_matched %>% 
              select(LakeID, trend, season, Variable) %>%
              arrange(LakeID) %>%
              pivot_wider(names_from = Variable,
                          values_from = trend))  %>%
  pivot_wider(names_from = season,
              values_from = Air_temp_C:Shortwave_mJm2) %>%
  left_join(mem %>% select(M, LakeID, var) %>% filter(var == "summer_bot_mean")) %>%
  mutate(M_binary = ifelse(M > 0, "> 0", "<= 0"))

responses <- c("summer_bot_mean", "M_binary")
potential_drivers <- c("Air_temp_C_spring", "Air_temp_C_summer", "M")
mem_rf_df_bot_temp <- df_for_rf %>%
  select(all_of(c(responses, potential_drivers))) %>%
  na.omit() %>%
  mutate(target = "Bottom-water\ntemperature")
ns <- mem_rf_df_bot_temp %>%
  group_by(M_binary) %>%
  summarize(n = n())
rf1_bot <- randomForest(summer_bot_mean ~ Air_temp_C_spring + Air_temp_C_summer,
                    data = mem_rf_df_bot_temp %>%
                      filter(M_binary == "> 0"), 
                    ntree = 1000, 
                    importance = T)
rf2_bot <- randomForest(summer_bot_mean ~ Air_temp_C_spring + Air_temp_C_summer,
                    data = mem_rf_df_bot_temp %>%
                      filter(M_binary == "<= 0"), 
                    ntree = 1000, 
                    importance = T)
ImpData_bot_temp <- as.data.frame(importance(rf1_bot)) %>%
  mutate(Var.Names = row.names(.)) %>%
  mutate(target = "Bottom-water\ntemperature",
         M_binary = "> 0") %>%
  rbind(rbind(as.data.frame(importance(rf2_bot)) %>%
               mutate(Var.Names = row.names(.)) %>%
               mutate(target = "Bottom-water\ntemperature",
                      M_binary = "<= 0"))) %>%
  left_join(ns)

mem_rf_df_bot_temp %>%
  ggplot(aes(x = Air_temp_C_spring, y = summer_bot_mean)) +
  geom_point()+
  facet_wrap(~M_binary) 

partial_bot_summer <- partialPlot(rf1_bot, as.data.frame(mem_rf_df_bot_temp), "Air_temp_C_summer")
partial_bot_spring <- partialPlot(rf1_bot, as.data.frame(mem_rf_df_bot_temp), "Air_temp_C_spring")
```

Repeat for DO

```{r}
met_trends_matched <- max_cor_dates %>%
  filter(var == "Summer hypo. DO") %>%
  left_join(met_by_bot_do, 
            by = c("LakeID", "max_doy" = "doy"), 
            relationship = "many-to-many") %>%
  group_by(season, var) %>%
  rename(cor_date_var = var)

df_for_rf <- summer_bot_do %>%
  left_join(met_trends_matched %>% 
              select(LakeID, trend, season, Variable) %>%
              arrange(LakeID) %>%
              pivot_wider(names_from = Variable,
                          values_from = trend))  %>%
  pivot_wider(names_from = season,
              values_from = Air_temp_C:Shortwave_mJm2) %>%
  left_join(mem %>% select(M, LakeID, var) %>% filter(var == "summer_bot_do")) %>%
  left_join(summer_surf_tp) %>%
  mutate(M_binary = ifelse(M > 0, "> 0", "<= 0"))

responses <- c("summer_bot_do", "M_binary")
potential_drivers <- c("Air_temp_C_spring", "Air_temp_C_summer")
mem_rf_df_bot_do <- df_for_rf %>%
  select(all_of(c(responses, potential_drivers))) %>%
  na.omit() %>%
  mutate(target = "Bottom-water\ndissolved oxygen")
ns <- mem_rf_df_bot_do %>%
  group_by(M_binary) %>%
  summarize(n = n())
rf1 <- randomForest(summer_bot_do ~ Air_temp_C_spring + Air_temp_C_summer,
                    data = mem_rf_df_bot_do %>%
                      filter(M_binary == "> 0"),, 
                    ntree = 1000, 
                    importance = T)
rf2 <- randomForest(summer_bot_do ~ Air_temp_C_spring + Air_temp_C_summer,
                    data = mem_rf_df_bot_do %>%
                      filter(M_binary == "<= 0"), 
                    ntree = 1000, 
                    importance = T)

ImpData_bot_DO <- as.data.frame(importance(rf1)) %>%
  mutate(Var.Names = row.names(.)) %>%
  mutate(target = "Bottom-water\nDO",
         M_binary = "> 0") %>%
  rbind(rbind(as.data.frame(importance(rf2)) %>%
               mutate(Var.Names = row.names(.)) %>%
               mutate(target = "Bottom-water\nDO",
                      M_binary = "<= 0"))) %>%
  left_join(ns)

ImpData_bot_DO %>%
  ggplot(aes(x = `%IncMSE`, y = Var.Names)) +
  geom_col() +
  theme_bw() +
  facet_wrap(~target)

partial_do_summer <- partialPlot(rf1, as.data.frame(mem_rf_df_bot_do), "Air_temp_C_summer")
partial_do_spring <- partialPlot(rf1, as.data.frame(mem_rf_df_bot_do), "Air_temp_C_spring")
```

Surface temp

```{r}
met_trends_matched <- max_cor_dates %>%
  filter(var == "Summer epi. temperature") %>%
  left_join(met_by_surf_temp, 
            by = c("LakeID", "max_doy" = "doy"), 
            relationship = "many-to-many") %>%
  group_by(season, var) %>%
  rename(cor_date_var = var)

df_for_rf <- summer_surf_mean %>%
  left_join(met_trends_matched %>% 
              select(LakeID, trend, season, Variable) %>%
              arrange(LakeID) %>%
              pivot_wider(names_from = Variable,
                          values_from = trend))  %>%
  pivot_wider(names_from = season,
              values_from = Air_temp_C:Shortwave_mJm2) %>%
  left_join(mem %>% select(M, LakeID, var) %>% filter(var == "summer_surf_mean")) %>%
  mutate(M_binary = ifelse(M > 0, "> 0", "<= 0"))

responses <- c("summer_surf_mean", "M_binary")
potential_drivers <- c("Air_temp_C_spring", "Air_temp_C_summer")
mem_rf_df_surf_temp <- df_for_rf %>%
  select(all_of(c(responses, potential_drivers))) %>%
  na.omit() %>%
  mutate(target = "Surface-water\ntemperature")

rf1 <- randomForest(summer_surf_mean ~ Air_temp_C_spring + Air_temp_C_summer,
                    data = mem_rf_df_surf_temp %>%
                      filter(M_binary == "> 0"),, 
                    ntree = 1000, 
                    importance = T)
rf2 <- randomForest(summer_surf_mean ~ Air_temp_C_spring + Air_temp_C_summer,
                    data = mem_rf_df_surf_temp %>%
                      filter(M_binary == "<= 0"), 
                    ntree = 1000, 
                    importance = T)
ns <- mem_rf_df_surf_temp %>%
  group_by(M_binary) %>%
  summarize(n = n())
ImpData_surf_temp <- as.data.frame(importance(rf1)) %>%
  mutate(Var.Names = row.names(.)) %>%
  mutate(target = "Surface-water\ntemperature",
         M_binary = "> 0") %>%
  rbind(as.data.frame(importance(rf2)) %>%
          mutate(Var.Names = row.names(.)) %>%
          mutate(target = "Surface-water\ntemperature",
                 M_binary = "<= 0")) %>%
  left_join(ns)

partial_surf_summer <- partialPlot(rf1, as.data.frame(mem_rf_df_surf_temp), "Air_temp_C_summer")
partial_surf_spring <- partialPlot(rf1, as.data.frame(mem_rf_df_surf_temp), "Air_temp_C_spring")
```

Combine and plot

```{r}
ImpData_bot_temp %>%
  rbind(ImpData_bot_DO) %>%
  rbind(ImpData_surf_temp) %>%
  mutate(neg = `%IncMSE` > 0) %>%
  mutate(Var.Names = factor(Var.Names, levels = c("Air_temp_C_spring",
                                                  "Air_temp_C_summer"),
                            labels = c("Spring", "Summer")),
         M_binary = factor(M_binary,
                           levels = c("> 0", "<= 0"),
                           labels = c("SEM > 0", "SEM <= 0"))) %>%
  ggplot(aes(y = `%IncMSE`, x = Var.Names, fill = Var.Names)) +
  geom_hline(yintercept = 0, color = "grey70") +
  geom_col(aes(alpha = neg)) +
  theme_bw() +
  ylab("Variable importance\n(% increase in MSE)") +
  xlab("Air temperature trend") +
  scale_fill_manual(values = c("Spring" = "#25469a",
                               "Summer" = "#c16586"))+
  scale_alpha_manual(values = c(0.2, 1)) +
  facet_grid(target~M_binary, switch = "y") +
  theme(
    strip.placement = "outside",   # format to look like title
    strip.background = element_blank(),
    #strip.text.y = element_blank(),
    strip.clip = "off",
    panel.spacing.y = unit(0.8, "lines"),
    axis.text.x = element_text(vjust=-0.5),
    axis.title.y = element_text(size = 9),
    legend.position = "none"
  )
```

```{r}
imp <- ImpData_bot_temp %>%
  mutate(Var.Names = factor(Var.Names, levels = c("Air_temp_C_spring",
                                                  "Air_temp_C_summer"),
                            labels = c("Spring", "Summer")),
         M_binary = factor(M_binary,
                           levels = c("> 0", "<= 0"),
                           labels = c("SEM > 0", "SEM <= 0"))) %>%
  mutate(neg = `%IncMSE` > 0,
         target = "") %>%
  ggplot(aes(y = `%IncMSE`, x = Var.Names, fill = Var.Names)) +
  geom_hline(yintercept = 0, color = "grey70") +
  geom_col(aes(alpha = neg)) +
  theme_bw() +
  ylab("Variable importance\n(% increase in MSE)") +
  xlab("Air temperature trend") +
  scale_fill_manual(values = c("Spring" = "#25469a",
                               "Summer" = "#c16586"))+
  scale_alpha_manual(values = c(0.2, 1)) +
  facet_grid(M_binary~target) +
  theme(
    strip.placement = "outside",   # format to look like title
    strip.background = element_blank(),
    #strip.text.y = element_blank(),
    strip.clip = "off",
    panel.spacing.y = unit(0.8, "lines"),
    axis.text.x = element_text(vjust=-0.5),
    axis.title.y = element_text(size = 9),
    legend.position = "none"
  )

partial_bot_summer_greater <- data.frame(
  partialPlot(rf1_bot, as.data.frame(mem_rf_df_bot_temp), "Air_temp_C_summer")) %>%
    mutate(var = "Summer")
partial_bot_spring_greater <- data.frame(
  partialPlot(rf1_bot, as.data.frame(mem_rf_df_bot_temp), "Air_temp_C_spring")) %>%
    mutate(var = "Spring")
partial_bot_summer_less <- data.frame(
  partialPlot(rf2_bot, as.data.frame(mem_rf_df_bot_temp), "Air_temp_C_summer")) %>%
    mutate(var = "Summer")
partial_bot_spring_less <- data.frame(
  partialPlot(rf2_bot, as.data.frame(mem_rf_df_bot_temp), "Air_temp_C_spring")) %>%
    mutate(var = "Spring")

M_part <- partial_bot_summer_greater %>%
  full_join(partial_bot_spring_greater) %>%
  mutate(M_binary = "SEM > 0") %>%
  full_join(partial_bot_summer_less %>%
              full_join(partial_bot_spring_less) %>%
              mutate(M_binary = "SEM <= 0"))

rug_data <- mem_rf_df_bot_temp %>%
  pivot_longer(cols = all_of(potential_drivers),
               names_to = "var", values_to = "x") %>%
  filter(!is.na(x)) %>%
  mutate(var = factor(var, levels = c("Air_temp_C_spring",
                                                  "Air_temp_C_summer"),
                            labels = c("Spring", "Summer")),
         M_binary = factor(M_binary,
                           levels = c("> 0", "<= 0"),
                           labels = c("SEM > 0", "SEM <= 0")))

part <- M_part %>%
  mutate(M_binary = factor(M_binary,
                           levels = c("SEM > 0", "SEM <= 0"),
                           labels = c("SEM > 0", "SEM <= 0"))) %>%
  ggplot(aes(x = x)) +
  geom_hline(yintercept = 0, color = "grey60", linewidth = 0.3) +
  geom_vline(xintercept = 0, color = "grey60", linewidth = 0.3) +
  geom_line(aes(y = y)) +
  geom_rug(aes(x = x), sides = "b", data = rug_data, alpha = 0.4, outside = TRUE, length = unit(0.08, "npc")) +
  coord_cartesian(clip = "off") +
  facet_grid(M_binary~var, scales = "free_x") +
  xlab("Air temperature trend") +
  theme_bw() +
  theme(
    strip.placement = "outside",   # format to look like title
    strip.background = element_blank(),
    #strip.text.y = element_blank(),
    strip.clip = "off",
    panel.spacing.y = unit(0.8, "lines"),
    axis.text.x = element_text(vjust=-0.5),
    axis.title.y = element_text(size = 9),
    legend.position = "none"
  ) +
  ylab("Bottom-water temperature trend")

jpeg("../Figures/Bottom-water temperature trend drivers.jpeg", width = 6, height = 3, units = "in", res = 300)
ggarrange(imp, part, widths = c(.5,1))
dev.off()
```


met_trends_matched <- max_cor_dates %>%
  filter(var == "Summer hypo. temperature") %>%
  left_join(met_by_bot_temp, 
            by = c("LakeID", "max_doy" = "doy"), 
            relationship = "many-to-many") %>%
  group_by(season, var) %>%
  rename(cor_date_var = var)

df_for_rf <- summer_bot_mean %>%
  left_join(met_trends_matched %>% 
              select(LakeID, trend, season, Variable) %>%
              arrange(LakeID) %>%
              pivot_wider(names_from = Variable,
                          values_from = trend))  %>%
  pivot_wider(names_from = season,
              values_from = Air_temp_C:Shortwave_mJm2) %>%
  left_join(mem %>% select(M, LakeID, var) %>% filter(var == "summer_bot_mean")) %>%
  mutate(M_binary = ifelse(M > 0, "> 0", "<= 0"))

responses <- c("summer_bot_mean", "M_binary")
potential_drivers <- c("Air_temp_C_spring", "Air_temp_C_summer", "M")
mem_rf_df_bot_temp <- df_for_rf %>%
  select(all_of(c(responses, potential_drivers))) %>%
  na.omit() %>%
  mutate(target = "Surface-water\ntemperature")

rf1 <- randomForest(summer_bot_mean ~ Air_temp_C_spring + Air_temp_C_summer + M,
                    data = mem_rf_df_bot_temp, 
                    ntree = 1000, 
                    importance = T)

library(vivid)
viviRF <- vivi(rf1, 
               data = mem_rf_df_bot_temp,
               response = "summer_bot_mean",
               gridSize = 50,
               importanceType = "agnostic",
               nmax = 500,
               reorder = TRUE,
               predictFun = NULL,
               numPerm = 4,
               showVimpError = FALSE)
viviNetwork(mat = viviRF)
top5 <- colnames(viviRF)[1:3]
pdpVars(data = mem_rf_df_bot_temp,
        fit = rf1,
        response = 'summer_bot_mean',
        vars = top5,
        nIce = 100)

pdpPairs(data = mem_rf_df_bot_temp,
         fit = rf1,
         response = 'summer_bot_mean',
         nmax = 500, 
         gridSize = 10,         
         vars = colnames(viviRF)[1:3],
         nIce = 100)

p <- pdpZen(data = mem_rf_df_bot_temp, 
       fit = rf1, 
       response = 'summer_bot_mean', 
       nmax = 500, 
       zpath = colnames(viviRF)[1:3],
       gridSize = 10)

#+
  geom_hline(yintercept = 0, color = "grey70")
```


M_part <- partials_surf_temp %>%
  full_join(partials_hypo_temp) %>%
  full_join(partials_hypo_do) %>%
  full_join(partials_vhod)  %>%
  mutate(Type = "M")

part <- M_part %>%
  mutate(target = factor(target, levels = c("Surface-water\ntemperature",
                                            "Bottom-water\ntemperature", 
                                            "Bottom-water\noxygen demand",
                                            "Bottom-water\ndissolved oxygen"))) %>%
  group_by(Type, var, target) %>%
  arrange(x, .by_group = TRUE) %>%
  reframe(x_interp = approx(x = x, y = y, xout = seq(min(x), max(x), length.out = 10000))$x,
          y_interp = approx(x = x, y = y, xout = seq(min(x), max(x), length.out = 10000))$y) %>%
  ggplot(aes(x = x_interp)) +
  geom_hline(yintercept = 0, color = "grey60", linewidth = 0.3) +
  geom_point(aes(y = y_interp, color = y_interp), size = 0.1) +
  geom_rug(aes(x = x), sides = "b", data = rug_data, alpha = 0.4, outside = TRUE, length = unit(0.08, "npc")) +
  coord_cartesian(clip = "off") +
  facet_grid(target~var, scales = "free_x", switch = "both") +
  labs(x=NULL) +
  theme_bw() +
  scale_color_gradientn(colors = c("#25469a", "#735690","#c16586"), 
                        limits = c(-0.3, 0.3)) +
  theme(
    strip.placement = "outside",   # format to look like title
    strip.background = element_blank(),
    #strip.text.y = element_blank(),
    strip.clip = "off",
    panel.spacing.y = unit(0.8, "lines"),
    axis.text.x = element_text(vjust=-0.5),
    axis.title.y = element_text(size = 9),
    legend.position = "none"
  ) +
  ylab("Partial dependence of memory statistic")
```

