---
title: "Air temp analysis"
author: "Abby Lewis"
date: "2023-05-22"
output: html_document
---

This file analyzes the effect of variation in air temperature on hypolimnetic temperature and DO. 

Table of contents:
Step 1: Load packages and data
Step 2: Generate temp anomaly summary fig
Step 3: Analyze the effect of monthly air temperature on hypolimnetic DO, VHOD, and temp dynamics


Step 1: Load packages and data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(mosaic)
library(ggallin)
library(ppcor)
library(multcompView)
library(ggpubr)
library(viridis)
library(MASS)
source("monthly_correlations.R")

#Load saved data
with_temp = read.csv("../Compiled data/All_data_annual AF.csv")

# Make a directory for Figures if there isn't one
if (!dir.exists("../Figures")){dir.create("../Figs")}
```


Step 2: Generate temp anomaly summary fig
```{r}
zscore <- function(data, na.rm = T){(data-mean(data, na.rm = na.rm))/sd(data, na.rm = na.rm)}
do_anomaly_stat <- with_temp%>%
  group_by(LakeID)%>%
  mutate(hypo_do_anom = DO_mgL_HYPO-mean(DO_mgL_HYPO, na.rm = T),
         spr_temp = (temp_mar+temp_apr)/2,
         spr_temp_anom = spr_temp - mean(spr_temp, na.rm = T))

cor_data <- do_anomaly_stat%>%
  dplyr::select(hypo_do_anom,spr_temp_anom,Year)%>%
  na.omit()

length(unique(cor_data$LakeID))

pcor.test(cor_data$hypo_do_anom, cor_data$spr_temp_anom,cor_data$Year,method = "spearman")$estimate

do_air_sum <- do_anomaly_stat%>%
  filter(!is.na(spr_temp_anom),
         !is.na(hypo_do_anom))%>%
  group_by(LakeID)%>%
  mutate(hypoxic = ifelse(max(DO_mgL_HYPO,na.rm = T)<1,"y","n"))%>%
  ungroup()%>%
  filter(hypoxic=="n")%>%
  mutate(temp_anom_sum = ifelse(spr_temp_anom>0.6,"> 0.6",
                                ifelse(spr_temp_anom<(-0.6),"< -0.6","between -0.6 \nand 0.6")),
         temp_anom_sum = factor(temp_anom_sum, levels = c("< -0.6", "between -0.6 \nand 0.6", "> 0.6")),
         do_anom_sum = ifelse(hypo_do_anom>0.2,"> 0.2",
                                ifelse(hypo_do_anom<(-0.2),"< -0.2","between -0.2 \nand 0.2")),
         do_anom_sum = factor(do_anom_sum, levels = c("> 0.2", "between -0.2 \nand 0.2", "< -0.2")))%>%
  group_by(temp_anom_sum)%>%
  mutate(n=n())

jpeg("../Figures/DO and air temp anomaly - spring.jpg", width = 5, height = 3, res = 300, units = "in")
do_air_sum%>%
  ggplot(aes(fill = do_anom_sum, x = temp_anom_sum))+
  geom_bar(position = "fill")+
  geom_hline(yintercept = .5, color = "grey80",lty = "dashed")+
  geom_text(aes(y = .9, label = paste0("n = ",n)), color = "grey50")+
  labs(fill="Hypolimnetic \nDO anomaly (mg/L)")+
  xlab("Spring air temperature anomaly (ºC)")+
  ylab("Percent of lake-years")+
  scale_y_continuous(labels = scales::percent)+
  theme_bw()+
  scale_fill_viridis(discrete=T, direction = -1)
dev.off()
```


Step 3: Analyze the effect of monthly air temperature on hypolimnetic DO, VHOD, and temp dynamics
```{r}
many_lake_temp_sum <- with_temp%>%
  pivot_longer(cols = contains("temp", ignore.case = F), names_to = "mon")%>%
  filter(!mon=="mean_temp")%>%
  mutate(mon = month(as.Date(paste0("2022-",substr(mon,6,8),"-01"), format = "%Y-%b-%d")),
         Year = ifelse(mon>=9,Year+1,Year))
# ---

jpeg("../Figures/DO and temp by month.jpg", width = 6, height = 4, res = 300, units = "in")
do_temp <- many_lake_temp_sum %>%
  mutate(hypoxic = ifelse(max(DO_mgL_HYPO, na.rm = T)<1,"yes","no"))%>%
  filter(hypoxic=="no") %>%
  monthly_correlations("DO_mgL_HYPO", "value", "hypolimnetic DO", alpha = 0.001)
do_temp
dev.off()

jpeg("../Figures/AF and temp by month.jpg", width = 6, height = 4, res = 300, units = "in")
do_temp_af <- many_lake_temp_sum %>%
  mutate(oxic = ifelse(max(AF, na.rm = T) == 0, "yes", "no"))%>%
  filter(oxic == "no") %>%
  monthly_correlations("AF", "value", "anoxic factor", alpha = 0.001)
do_temp_af
dev.off()

jpeg("../Figures/Air and strat hypo temp by month.jpg", width = 6, height = 4, res = 300, units = "in")
temp_temp <- monthly_correlations(many_lake_temp_sum,
                                  "strat_Temp_C_HYPO", "value", 
                                  "hypolimnetic temperature", alpha = 0.001)
temp_temp
dev.off()

jpeg("../Figures/Demand and air temp by month.jpg", width = 6, height = 4, res = 300, units = "in")
temp_demand <- monthly_correlations(many_lake_temp_sum, 
                                    "DO_demand_mgLd_HYPO", "value", 
                                    "VHOD", alpha = 0.001)
temp_demand
dev.off()

#Quick check to see if this shows up for TP --> not really
jpeg("../Figures/TP and air temp by month.jpg", width = 6, height = 4, res = 300, units = "in")
temp_tp <- monthly_correlations(many_lake_temp_sum, 
                                    "strat_TP_ugL_HYPO", "value", 
                                    "hypolimnetic TP", alpha = 0.05)
temp_tp
dev.off()

jpeg("../Figures/Monthly correlations.jpg", width = 5, height = 8, res = 300, units = "in")
ggarrange(temp_temp+theme(axis.title.x = element_blank()),
          temp_demand+theme(axis.title.x = element_blank()),
          do_temp,
          common.legend = T, legend = "bottom",nrow=3, labels = "auto", align = "hv")
dev.off()

jpeg("../Figures/Monthly correlations af.jpg", width = 5, height = 9, res = 300, units = "in")
ggarrange(temp_temp + theme(axis.title.x = element_blank()),
          temp_demand + theme(axis.title.x = element_blank()),
          do_temp + theme(axis.title.x = element_blank()),
          do_temp_af,
          common.legend = T, legend = "bottom", nrow=4, 
          labels = "auto", align = "hv")
dev.off()
```


Epi and hypo temp and DO
```{r}
jpeg("../Figures/DO and temp by month - EPI.jpg", width = 6, height = 4, res = 300, units = "in")
do_temp_epi <- many_lake_temp_sum %>%
  mutate(hypoxic = ifelse(max(DO_mgL_EPI, na.rm = T)<1,"yes","no"))%>%
  filter(hypoxic=="no") %>%
  monthly_correlations("DO_mgL_EPI", "value", "epilimnetic DO", alpha = 0.001)
do_temp_epi
dev.off()

jpeg("../Figures/Air and epi temp by month.jpg", width = 6, height = 4, res = 300, units = "in")
temp_temp_epi <- monthly_correlations(many_lake_temp_sum,
                                  "Temp_C_EPI", "value", 
                                  "epilimnetic temperature", alpha = 0.001)
temp_temp_epi
dev.off()

jpeg("../Figures/Monthly correlations - EPI.jpg", width = 5, height = 5, res = 300, units = "in")
ggarrange(temp_temp_epi+theme(axis.title.x = element_blank()),
          do_temp_epi,
          common.legend = T, legend = "bottom",nrow=2, labels = "auto", align = "hv")
dev.off()

jpeg("../Figures/Monthly correlations - EPI vs HYPO.jpg", width = 5, height = 5, res = 300, units = "in")
ggarrange(temp_temp_epi + theme(axis.title.x = element_blank()),
          temp_temp,
          common.legend = T, legend = "bottom",nrow=2, labels = "auto", align = "hv")
dev.off()
```

AGU figures
```{r}
temp_temp_epi_agu <- monthly_correlations_talk(many_lake_temp_sum,
                                  "Temp_C_EPI", "value", 
                                  "surface-water temperature", alpha = 0.001)

temp_temp_agu <- monthly_correlations_talk(many_lake_temp_sum,
                                  "strat_Temp_C_HYPO", "value", 
                                  "bottom-water temperature", alpha = 0.001)

jpeg("../Figures/Monthly correlations - EPI vs HYPO AGU.jpg", 
     width = 5, height = 5, res = 300, units = "in")
ggarrange(temp_temp_epi_agu + theme(axis.title.x = element_blank()),
          temp_temp_agu,
          common.legend = T, legend = "bottom",nrow=2, align = "hv")
dev.off()

temp_temp_epi_agu_aug <- monthly_correlations_talk_aug(many_lake_temp_sum,
                                  "Temp_C_EPI", "value", 
                                  "surface-water temperature", alpha = 0.001)

temp_temp_agu_aug <- monthly_correlations_talk_aug(many_lake_temp_sum,
                                  "strat_Temp_C_HYPO", "value", 
                                  "bottom-water temperature", alpha = 0.001)

temp_temp_agu_spring_summer <- monthly_correlations_spring_summer(many_lake_temp_sum,
                                  "strat_Temp_C_HYPO", "value", 
                                  "bottom-water temperature", alpha = 0.001)

temp_temp_epi_agu_spring_summer <- monthly_correlations_spring_summer(many_lake_temp_sum,
                                  "Temp_C_EPI", "value", 
                                  "bottom-water temperature", alpha = 0.001)

jpeg("../Figures/Monthly correlations - EPI vs HYPO AGU aug.jpg", 
     width = 5, height = 5, res = 300, units = "in")
ggarrange(temp_temp_epi_agu_aug + theme(axis.title.x = element_blank()),
          temp_temp_agu_aug,
          common.legend = T, legend = "bottom",nrow=2, align = "hv")
dev.off()

jpeg("../Figures/Monthly correlations - EPI vs HYPO AGU aug hypo.jpg", 
     width = 5, height = 5, res = 300, units = "in")
ggarrange(temp_temp_epi_agu + theme(axis.title.x = element_blank()),
          temp_temp_agu_aug,
          common.legend = T, legend = "bottom",nrow=2, align = "hv")
dev.off()

jpeg("../Figures/Monthly correlations - EPI vs HYPO AGU spring summer.jpg", 
     width = 5, height = 5, res = 300, units = "in")
ggarrange(temp_temp_epi_agu_spring_summer + 
            theme(axis.title.x = element_blank()) +
            ggtitle("Summer surface-water temperature"),
          temp_temp_agu_spring_summer +
            ggtitle("Summer bottom-water temperature"),
          common.legend = T, legend = "right",nrow=2, align = "hv")
dev.off()

do_temp_agu <- many_lake_temp_sum %>%
  mutate(hypoxic = ifelse(max(DO_mgL_HYPO, na.rm = T)<1,"yes","no"))%>%
  filter(hypoxic=="no") %>%
  monthly_correlations_talk("DO_mgL_HYPO", "value", "bottom-water DO", alpha = 0.001)

temp_demand_agu <- monthly_correlations_talk(many_lake_temp_sum, 
                                    "DO_demand_mgLd_HYPO", "value", 
                                    "oxygen demand", alpha = 0.001)

do_temp_agu_spring_summer <- many_lake_temp_sum %>%
  mutate(hypoxic = ifelse(max(DO_mgL_HYPO, na.rm = T)<1,"yes","no"))%>%
  filter(hypoxic=="no") %>%
  monthly_correlations_spring_summer("DO_mgL_HYPO", "value", 
                                  "bottom-water DO", alpha = 0.001)

temp_demand_agu_spring_summer <- monthly_correlations_spring_summer(many_lake_temp_sum,
                                  "DO_demand_mgLd_HYPO", "value", 
                                  "oxygen demand", alpha = 0.001)

jpeg("../Figures/Monthly correlations - VHOD and DO AGU.jpg", width = 5, height = 5, res = 300, units = "in")
ggarrange(temp_demand_agu + theme(axis.title.x = element_blank()),
          do_temp_agu,
          common.legend = T, legend = "bottom",nrow=2, align = "hv")
dev.off()

jpeg("../Figures/Monthly correlations - VHOD and DO AGU spring summer.jpg", 
     width = 5, height = 5, res = 300, units = "in")
ggarrange(temp_demand_agu_spring_summer + 
            theme(axis.title.x = element_blank()) +
            ggtitle("Summer bottom-water oxygen demand"),
          do_temp_agu_spring_summer +
            ggtitle("Summer bottom-water dissolved oxygen"),
          common.legend = T, legend = "right",nrow=2, align = "hv")
dev.off()

do_temp_agu_aug <- many_lake_temp_sum %>%
  mutate(hypoxic = ifelse(max(DO_mgL_HYPO, na.rm = T)<1,"yes","no"))%>%
  filter(hypoxic=="no") %>%
  monthly_correlations_talk_aug("DO_mgL_HYPO", "value", "bottom-water DO", alpha = 0.001)

temp_demand_agu_aug <- monthly_correlations_talk_aug(many_lake_temp_sum, 
                                    "DO_demand_mgLd_HYPO", "value", 
                                    "oxygen demand", alpha = 0.001)

jpeg("../Figures/Monthly correlations - VHOD and DO AGU aug.jpg", width = 5, height = 5, res = 300, units = "in")
ggarrange(temp_demand_agu_aug + theme(axis.title.x = element_blank()),
          do_temp_agu_aug,
          common.legend = T, legend = "bottom",nrow=2, align = "hv")
dev.off()
```


```{r}
library(ggthemes)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
#remotes::install_github("ropensci/rnaturalearthhires")
library(rnaturalearthhires)
library(ggspatial)
library(viridis)
library(ggpubr)

many_lake_stat <- many_lake_temp_sum%>%
  filter(!is.na(strat_Temp_C_HYPO),!is.na(mon), !is.na(value))%>%
  group_by(LakeID, mon)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10,
         #max(value)-min(value)>2
         )%>%
  dplyr::summarize(temp_temp = pcor.test(strat_Temp_C_HYPO, value, Year, method = "spearman")$estimate,
            )

lat_long = read.csv("https://pasta-s.lternet.edu/package/data/eml/edi/1029/9/fadd3eaa25b5fdd1fc4efba70e660579")

to_plot <- many_lake_stat %>%
  filter(mon < 8) %>%
  group_by(LakeID)%>%
  summarize(best_mon = mon[which.max(temp_temp)]) %>%
  left_join(lat_long) 

library(ggridges)

to_plot %>%
  ggplot(aes(x = Latitude_DD, y = best_mon)) +
  geom_point()

to_plot %>%
  mutate(rounded_lat = round(Latitude_DD,0),
         Region = ifelse(StateOrProvince %in% c("Wisconsin", "Minnesota", "Michigan"),
                         "Midwest",
                         ifelse(StateOrProvince %in% c("Maine", "New York", "NH", "New Hampshire", "Vermont"), "NE", NA))) %>%
  group_by(rounded_lat, Region) %>%
  mutate(n = n()) %>%
  filter(n > 10,
         Country == "USA",
         !is.na(Region)) %>%
  ggplot(aes(y = as.factor(rounded_lat), x = best_mon)) +
  geom_density_ridges() +
  facet_wrap(~Region)
```

Map
```{r}
for_states <- to_plot
states <- ne_states(returnclass = "sf",country = "United States of America")

us_map = ggplot(data = states) +
  geom_sf() +
  coord_sf(expand = FALSE, ylim = c(42, 47.5), xlim = c(-93,-86.5))+
  geom_point(data = for_states, 
             aes(Longitude_DD, Latitude_DD, fill = as.factor(best_mon)),
             shape = 21, color = "white", size = 2.8, alpha  =.5, stroke = .4)+
  theme_bw()+
  theme(plot.margin = margin(0, 0, 0, 0, "cm"),
    legend.box.background = element_rect(fill = "white", color = "white"))+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())+
  geom_hline(yintercept = 44)
us_map
  
us_map = ggplot(data = states) +
  geom_sf() +
  coord_sf(expand = FALSE, ylim = c(42, 47.5), xlim = c(-76,-67)) +
  geom_point(data = for_states, 
             aes(Longitude_DD, Latitude_DD, fill = as.factor(best_mon)),
             shape = 21, color = "white", size = 2.8, alpha  =.5, stroke = .4)+
  theme_bw()+
  theme(plot.margin = margin(0, 0, 0, 0, "cm"),
    legend.box.background = element_rect(fill = "white", color = "white"))+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  geom_hline(yintercept = 44)
us_map

world <- ne_countries(scale = "medium", returnclass = "sf")
for_euro <- to_plot

us_map = ggplot(data = world) +
  geom_sf() +
  coord_sf(expand = FALSE, ylim = c(43, 70), xlim = c(-5,30)) +
  geom_point(data = for_euro, 
             aes(Longitude_DD, Latitude_DD, fill = as.factor(best_mon)),
             shape = 21, color = "white", size = 2.8, alpha  =.5, stroke = .4)+
  theme_bw()+
  theme(plot.margin = margin(0, 0, 0, 0, "cm"),
    legend.box.background = element_rect(fill = "white", color = "white"))+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
us_map
```

```{r}
temp_temp_data <- many_lake_temp_sum%>%
  group_by(LakeID, mon)%>%
  filter(!is.na(strat_Temp_C_HYPO),!is.na(mon), !is.na(value))%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10,
         #max(value)-min(value)>2
         )%>%
  dplyr::summarize(temp_temp = pcor.test(strat_Temp_C_HYPO,value,Year,method = "spearman")$estimate,
            )

strat_avgs = read.csv("../Compiled data/stratified_averages.csv")%>%
  group_by(LakeID) %>%
  summarize(buoyancy_freq = median(buoyancy_freq, na.rm = T))

temp_temp_data %>% 
  left_join(strat_avgs) %>%
  pivot_wider(names_from = mon, values_from = temp_temp, names_prefix = "Month_") %>%
  ggplot(aes(y = Month_4 - Month_8, x = buoyancy_freq)) +
  geom_point() +
  geom_smooth() +
  ggpmisc::stat_poly_eq(ggpmisc::use_label(c("eq", "p")), coef.digits = 2) +
  ylab("Relative correlation between hypo. temp\nand spring vs summer air temp")
```

```{r}
lat_long <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/1530/1/fadd3eaa25b5fdd1fc4efba70e660579")
lat_long = lat_long %>%
  mutate(Region = ifelse(StateOrProvince %in% c("Wisconsin", "Minnesota"), "Midwest", NA),
         Region = ifelse(StateOrProvince %in% c("New York", "NY", "Maine", 
                                                "New Hampshire", "NH", "Vermont"), 
                         "Northeast", Region),
         Region = ifelse(Country == "Sweden", "Sweden", Region)) 

variable <- sym("Temp_C_HYPO")
value <- sym("value")
name <- "hypo. temp"
alpha = 0.001

many_lake_stat <- many_lake_temp_sum %>%
  filter(!is.na(!!variable), !is.na(mon), !is.na(!!value))%>%
  group_by(LakeID, mon)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10,
         #max(value)-min(value)>2
  )%>%
  dplyr::summarize(monthly_correlation = ppcor::pcor.test(!!variable, !!value, 
                                                          Year, method = "spearman")$estimate,
  )%>%
  filter(mon < 9) %>%
  left_join(lat_long) %>%
  filter(!is.na(Region))

labs = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug")

wilcox <- many_lake_stat%>%
  group_by(mon, Region)%>%
  mutate(p = wilcox.test(monthly_correlation)$p.value,
         n = n(),
         mon = factor(mon, levels = c(1:8),labels = labs)) %>%
  mutate(Region = paste0(Region, " (n = ",n,")"))
fig <- wilcox %>%
  ggplot(aes(x=mon, y = monthly_correlation))+
  geom_boxplot(aes(color=p<alpha))+
  #geom_text(aes(label = paste0("n = ",n), y = max(do_temp)))+ #n is the same across all
  geom_hline(yintercept=0)+
  ylim(-1,1.08)+
  xlab("Air temperature month")+
  ylab(paste0("Correlation with\n",name))+
  theme_bw()+
  scale_color_manual(values = c("grey50","#FB8B24"), breaks = c(F, T))+
  theme(axis.text.x = element_text(hjust = .5))+
  labs(color = paste0("p < ",alpha))+
  facet_wrap(~Region, ncol = 1)
jpeg("../Figures/Correlation with hypo. temp by region.jpeg", width = 6, height = 8, units = "in", res = 300)
fig
dev.off()
```

